---
title: "Data Cleaning - ZOVER outcome"
output: html_document
date: "2025-11-03"
---

**Manuscript title:** Ecological and socioeconomic factors associated with globally reported tick-borne viruses

**Corresponding author:** Samantha Sambado (sbsambado at ucsb.edu, ssambado at stanford.edu)

**Purpose**
To transform raw data from ZOVER database into variables that will be outcomes for the analysis which will be in the final file: zover_toprecognized_complete_pubmed_20240921.csv 


*Note: The ZOVER data is freely available under the Creative Commons Attribution-NonCommercial (CC BY-NC) license version 4.0 for personal and public non-commercial, research or academic use by individuals at academic, government or non-profit institutions. Users intending to use ZOVER data for commercial purposes should contact us via the official email.*

+ full ZOVER website (http://www.mgc.ac.cn/cgi-bin/ZOVER/main.cgi)
+ tick-virus website (http://www.mgc.ac.cn/cgi-bin/ZOVER/mainTable.cgi?db=tick)

Once on tick-virus wesbite, click on subgroups of tick-viruses (e.g. ssDNA), click "show all", click "save this table", open in personal excel, save as csv to Rproj

This data was accessed 2024-09-21


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

## upload necessary packages
library(readr)
library(tidyverse)
library(janitor)
library(lubridate)
library(stringr)
library(countrycode)
library(rnaturalearth)
library(easyPubMed)


## upload necessary data
# dsDNA_noRNA <- read_csv("tickviruses/raw/Tick-associated viruses (dsDNA viruses, no RNA stage) .csv", skip = 1) %>% clean_names()
# dsRNA <- read_csv("tickviruses/raw/Tick-associated viruses (dsRNA viruses) .csv", skip = 1) %>% clean_names()
# ssDNA <- read_csv("tickviruses/raw/Tick-associated viruses (ssDNA viruses) .csv", skip = 1) %>% clean_names()
# ssRNA_negstrands <- read_csv("tickviruses/raw/Tick-associated viruses (ssRNA negative-strand viruses) .csv", skip = 1) %>% clean_names()
# ssRNA_posstrands <- read_csv("tickviruses/raw/Tick-associated viruses (ssRNA positive-strand viruses, no DNA stage) .csv", skip = 1) %>% clean_names()
# unlcassified <- read_csv("tickviruses/raw/Tick-associated viruses (unclassified viruses) .csv", skip = 1) %>% clean_names()
```

#Step 1. Combine datasets
```{r}

## add unique identifier for each database
#dsDNA_noRNA$type <- "dsDNA_noRNA"
#dsRNA$type <- "dsRNA"
#ssDNA$type <- "ssDNA"
#ssRNA_negstrands$type <- "ssRNA_negstrands"
#ssRNA_posstrands$type <- "ssRNA_posstrands"
#unlcassified$type <- "unlcassified"
#
### combine all csv files into 1 csv file
#tick_viruses_raw <- rbind(dsDNA_noRNA, dsRNA, ssDNA, ssRNA_negstrands, ssRNA_posstrands, unlcassified #)
#
## add count column for EDA
#tick_viruses_raw$count <- 1
#
## check structure of dataframe
#str(tick_viruses_raw) # 8,110 x 11 (~500 more since accessed 2023-09-24)
```


#Step 2. Quick clean for EDA
```{r}
# clean up column names and data types
#tick_virus_all <- tick_viruses_raw %>% 
#  clean_names() %>% 
#  mutate_at(c('collection_year', 'sampling_country','type'),
#            as.factor) 
#
## check that levels are consistent
#length(unique(tick_virus_all$viruses)) # 7611
#length(unique(tick_virus_all$viral_family)) # 25
#length(unique(tick_virus_all$from_tick)) # 152
#length(unique(tick_virus_all$tick_genus)) # 13
#
## split up viruses into more manageable analysis chunks
#tick_virus_all[,12:13] <-str_split_fixed(tick_virus_all$viruses, c("virus|Virus|sp."), 2)
##tick_virus_all[,12:13] <- str_split_fixed(tick_virus_all$viruses, "virus", 2)
#names(tick_virus_all)[12] <- "virus_abbrev"
#names(tick_virus_all)[13] <- "virus_strain"
#
## check how much more this need works
## unique(tick_virus_all$virus_abbrev)
#
## trim extra white space at right side of label
#tick_virus_all$virus_abbrev <- str_trim(tick_virus_all$virus_abbrev,"right")
#
### lump some virus abbreviations together
## CCHF
#tick_virus_all$virus_abbrev[(tick_virus_all$virus_abbrev) =="Crimean-Congo hemorrhagic fever #orthonairo"] <- "Crimean-Congo hemorrhagic fever"
#
## Tick Phlebovirus
#tick_virus_all$virus_abbrev[(tick_virus_all$virus_abbrev) =="Phlebo"] <- "Tick phlebo"
#tick_virus_all$virus_abbrev[(tick_virus_all$virus_abbrev) =="unclassified Phlebo"] <- "Tick phlebo"
#tick_virus_all$virus_abbrev[(tick_virus_all$virus_abbrev) =="Jingmen"] <- "Jingmen tick"

```

#Step 3. Add spatial metadata
```{r}
# add iso3c code
#tick_virus_all$iso3c <- countrycode(tick_virus_all$sampling_country, 
#            origin = 'country.name',
#            destination = 'iso3c')
#
## add continent 
# tick_virus_all <-  tick_virus_all %>%
#   mutate(continent = countrycode(iso3c, origin = "iso3c",
#                                  destination = "continent"))
# 
## upload world map
#world <- ne_countries(scale = "large", returnclass = "sf")

```

#Step 4. More refined data cleaning

###Fix dates
The steps I decided to do:

1. observations with ~ = if it's same year fix to same year (ie 2019~2019 is just 2019)

2. observations with ~ = if it's across 2 years, go with the first (ie 2018-2019 is 2018)

3. see how many more observations with ~ = either omit or make a judgement call

4. separate years into every 5 years?

```{r}

#unique(tick_virus_all$collection_year)
#
### step 1. if same year fix to just 1 year
#tick_virus_all$collection_year[(tick_virus_all$collection_year) == "2022~2022"] <- "2022"
#tick_virus_all$collection_year[(tick_virus_all$collection_year) == "2020~2020"] <- "2020"
#tick_virus_all$collection_year[(tick_virus_all$collection_year) == "2019~2019"] <- "2019"
#tick_virus_all$collection_year[(tick_virus_all$collection_year) == "2017~2017"] <- "2017"
#tick_virus_all$collection_year[(tick_virus_all$collection_year) == "2016~2016"] <- "2016"
#tick_virus_all$collection_year[(tick_virus_all$collection_year) == "2015~2015"] <- "2015"
#tick_virus_all$collection_year[(tick_virus_all$collection_year) == "2013~2013"] <- "2013"
#tick_virus_all$collection_year[(tick_virus_all$collection_year) == "2005~2005"] <- "2005"
#
#
### step 2. if only 2 year span choose first year
#tick_virus_all$collection_year[(tick_virus_all$collection_year) == "2018~2019"] <- "2018"
#tick_virus_all$collection_year[(tick_virus_all$collection_year) == "2016~2017"] <- "2016"
#tick_virus_all$collection_year[(tick_virus_all$collection_year) == "2015~2016"] <- "2015"
#tick_virus_all$collection_year[(tick_virus_all$collection_year) == "2014~2015"] <- "2014"
#tick_virus_all$collection_year[(tick_virus_all$collection_year) == "2011~2012"] <- "2011"
#tick_virus_all$collection_year[(tick_virus_all$collection_year) == "2004~2005"] <- "2004"
#tick_virus_all$collection_year[(tick_virus_all$collection_year) == "2000~2001"] <- "2000"
#tick_virus_all$collection_year[(tick_virus_all$collection_year) == "1997~1998"] <- "1997"
#
### step 3a. how many other 1+ year strings are there?
#tick_virus_all %>% 
#  filter(grepl("~", collection_year)) # 123 observations
#
### step 3b. If it's only 2 year difference choose the middle year
#tick_virus_all$collection_year[(tick_virus_all$collection_year) == "2019~2021"] <- "2020"
#tick_virus_all$collection_year[(tick_virus_all$collection_year) == "2013~2015"] <- "2014"
#tick_virus_all$collection_year[(tick_virus_all$collection_year) == "2012~2014"] <- "2012"
#tick_virus_all$collection_year[(tick_virus_all$collection_year) == "2009~2011"] <- "2010"
#
### now there's 77 observations with years that are reported 2+ years together
#
#
### step 4. edit other random erros
#tick_virus_all$collection_year[(tick_virus_all$collection_year) == "0199"] <- "1990"
#
#tick_virus_all %>% 
#  filter(collection_year == "ND") # 462 years not reported
#
#
#### okay let's try to map this out as best as possible and make a decision about the 77 observations #with 2+ year aggregation reported
#
#unique(tick_virus_all$collection_year)
#
#tick_virus_all %>% 
#  filter(grepl("~", collection_year)) # 56 observations
#
#dim(tick_virus_all)# 8110 × 15
#
#
###### Aggregate observations by 5 years
#tick_virus_all <- tick_virus_all %>% 
#  mutate(collection_year = as.numeric(as.character(collection_year))) %>% 
#  filter(!(grepl("~", collection_year)) & collection_year != "ND") %>% 
#  mutate(year_agg = case_when(
#    collection_year >= 2020 ~ "2020",
#    collection_year >= 2015 ~ "2015",
#    collection_year >= 2010 ~ "2010",
#    collection_year >= 2005 ~ "2005",
#    collection_year >= 2000 ~ "2000",
#    collection_year >= 1995 ~ "1995",
#    collection_year >= 1990 ~ "1990",
#    collection_year >= 1985 ~ "1985",
#    collection_year >= 1980 ~ "1980",
#    collection_year >= 1975 ~ "1975",
#    collection_year >= 1970 ~ "1970",
#    collection_year >= 1965 ~ "1965",
#    collection_year >= 1960 ~ "1960",
#    collection_year >= 1955 ~ "1955",
#    collection_year >= 1950 ~ "1950",
#    collection_year < 1950 ~ "pre-1950"))
#
#tick_virus_all %>% 
#  filter(year_agg == "pre-1950") # only 3 left so I feel fine leaving it as that
```


###Check country names
```{r}
#unique(tick_virus_all$sampling_country) # 94/195 total
#
## let's see what's ND
#View(tick_virus_all %>%  # 28
#  filter(sampling_country == "ND") %>% 
#    group_by(references) %>% 
#    tally()) # 28 observations not listed.. checking reference
#
## many of these ND are cultures or unpublished.. maybe should get rid of both
#
#tick_virus_all %>% 
#  filter(sample_type == "Culture") #1,433 observations are cultured.. probably should remove these #but worry about that once I get to analysis

```

###Check viral family names
```{r}
# first check unique viral families
#unique(tick_virus_all$viral_family) # 23 unique families + 1 NA
#
#
## figure out what's NA
#tick_virus_all %>% 
#  filter(viral_family == "N/A") %>%  # 750 observations ... after 4 fixes its 467
#  group_by(sampling_country,viral_family) %>% 
#  tally() %>% # okay so a lot come from China, let's see if there's a common virus I can convert to #viral family
# arrange(desc(n))
#
## fixing NA from China
#tick_virus_all %>% 
#  filter(viral_family == "N/A" & sampling_country == "China") %>% 
#  group_by(virus_abbrev) %>% 
#  tally() %>%  # bole tick bole,  hyalomma asisticum, 
#  arrange(desc(n))
#
#tick_virus_all %>%
#  filter(grepl("Bole", viruses)) %>% 
#  group_by(viral_family) %>% 
#  tally() %>% 
#  arrange(desc(n))#  okay so bole viruses can be phenuiviridae or rhabdoviridae... 63 observations #are NA why is that
#
#tick_virus_all %>%
#  filter(grepl("Bole", viruses) & viral_family == "N/A") 
#
#tick_virus_all %>%
#  filter(grepl("Jingmen", viruses)) # looks like it's flaviviridae
#
#tick_virus_all %>%
#  filter(grepl("Wuhan", viruses)) # looks like it's Rhabdoviridae
#
## fixing NA from Australia
#tick_virus_all %>% 
#  filter(viral_family == "N/A" & sampling_country == "Australia") %>% 
#  group_by(virus_abbrev) %>% # so nothing in australia has viral family resolved
#  tally() %>%  
#  arrange(desc(n))
#
#tick_virus_all %>%
#  filter(grepl("Shelly headland", viruses)) 
#
## fixing NAs from Russia
#tick_virus_all %>% 
#  filter(viral_family == "N/A" & sampling_country == "Russia") %>% 
#  group_by(virus_abbrev)
#
#
### okay fixed some let's check main NAs left 467.. will try to par this down further later
#tick_virus_all %>% 
#  filter(viral_family == "N/A") %>%  # & sampling_country != "Australia") %>% 
#  group_by(virus_abbrev) %>% 
#  tally() %>% 
#  arrange(desc(n)) # bole tick 55
#
#
#
###### Edit viral family 
## first fix
#tick_virus_all <- 
#tick_virus_all %>%
#  mutate(viral_family = if_else(
#    grepl("Jingmen", viruses, fixed = TRUE),
#    "Flaviviridae", viral_family
#  )) 
#  # filter(grepl("Jingmen", viruses))
#
## second fix
#tick_virus_all <- 
#tick_virus_all %>%
#  mutate(viral_family = if_else(
#    grepl("Wuhan", viruses, fixed = TRUE),
#    "Rhabdoviridae", viral_family
#  )) #%>% 
#  #filter(grepl("Wuhan", viruses))
#
#
## 3rd fix
#tick_virus_all <- 
#  tick_virus_all %>%
#  mutate(viral_family = if_else(
#    grepl("Haseki", viruses, fixed = TRUE),
#    "Flaviviridae", viral_family
#  ))  
#  #filter(grepl("Haseki", viruses))
#
#
## 4rd fix
#tick_virus_all <- 
#tick_virus_all %>%
#  mutate(viral_family = if_else(
#    grepl("Yanggou", viruses, fixed = TRUE),
#    "Flaviviridae", viral_family
#  ))  
#  #filter(grepl("Yanggou", viruses))
```

###Make changes to individual virus names
(this is somewhat coarse - could be improved upon)
```{r}
# see all unique
#unique(tick_virus_all$virus_abbrev)
#
## pull out the most common ones
#
## test this function first
##tick_virus_all %>%
##  mutate(fix_viruses = case_when(grepl("Kemerovo virus", viruses) ~ "Kemerovo tickborne viral #fever")) %>% 
##  dplyr::select(c(2,18))
#
#
#tick_virus_all_cleanvirus <- tick_virus_all %>%
#  mutate(fix_viruses = case_when(grepl("African swine fever", viruses) ~ "African swine fever",
#                                 grepl("Colorado tick fever", viruses) ~ "Colorado tick fever",
#                                grepl("Lumpy skin disease", viruses) ~ "Lumpy skin disease",
#                                grepl("Kemerovo virus", viruses) ~ "Kemerovo tickborne viral fever",
#                                grepl("Pseudocowpox virus", viruses) ~ "Pseudocowpox virus",
#                                grepl("Eyach virus", viruses) ~ "Eyach virus", #TBE?
#                                grepl("Crimean-Congo hemorrhagic fever", viruses) ~ "Crimean-Congo #hemorrhagic fever",
#                                grepl("Okutama tick virus", viruses) ~ "Okutama tick virus",
#                                grepl("bole tick", viruses) ~ "Bole tick XXX",
#                                grepl("Bole Tick", viruses) ~ "Bole tick XXX",
#                                grepl("Bole tick", viruses) ~ "Bole tick XXX",
#                                grepl("Bocaparvovirus", viruses) ~ "Bocaparvovirus",
#                                grepl("Severe fever with thrombocytopenia virus", viruses) ~ "Severe #fever with thrombocytopenia virus",
#                                grepl("Nairobi sheep disease virus", viruses) ~ "Nairobi sheep #disease virus",
#                                grepl("Dabieshan Tick", viruses) ~ "Dabieshan Tick",
#                                grepl("St Croix River virus", viruses) ~ "St Croix River virus",
#                                grepl("Sawgrass", viruses) ~ "Sawgrass virus",
#                                grepl("Phlebovirus", viruses) ~ "Phlebovirus XXX",
#                                grepl("Taishun Tick Virus", viruses) ~ "Taishun Tick virus XXX",
#                                grepl("phlebovirus", viruses) ~ "Phlebovirus XXX",
#                                grepl("Wuhan Tick Virus 1", viruses) ~ "Wuhan tick virus 1",
#                                grepl("Wuhan Tick Virus 2", viruses) ~ "Wuhan tick virus 2",
#                                grepl("Wuhan tick virus 2", viruses) ~ "Wuhan tick virus 2",
#                                grepl("Tarumizu tick virus", viruses) ~ "Tarumizu tick virus",
#                                grepl("Guangdong tick orbivirus", viruses) ~ "Guangdong tick #orbivirus",
#                                grepl("Xinjiang tick orbivirus", viruses) ~ "Xinjiang tick #orbivirus",
#                                grepl("Okhotskiy virus", viruses) ~ "Okhotskiy virus",
#                                grepl("Huangpi Tick Virus", viruses) ~ "Huangpi tick Virus",
#                                grepl("Dugbe orthonairovirus ", viruses) ~ "Dugbe orthonairovirus #XXX",
#                                grepl("Tacheng Tick Virus 1", viruses) ~ "Tacheng tick virus 1 XXX",
#                                grepl("Yezo virus", viruses) ~ "Yezo virus XXX",
#                                grepl("Zhangzhou tick virus 1", viruses) ~ "Zhangzhou tick virus 1 #XXX",
#                                grepl("Songling virus", viruses) ~ "Songling virus XXX",
#                                grepl("Wad Medani virus", viruses) ~ "Wad Medani virus XXX",
#                                grepl("Tacheng Tick Virus ", viruses) ~ "Tacheng Tick virus XXX",    #                             
#                                grepl("Odaw virus", viruses) ~ "Odaw virus XXX",
#                                grepl("Guertu virus", viruses) ~ "Guertu virus XXX",
#                                grepl("Shibuyunji virus", viruses) ~ "Shibuyunji virus XXX",
#                                grepl("Quaranfil quaranjavirus", viruses) ~ "Quaranfil quaranjavirus #XXX",
#                                grepl("Tacheng Tick Virus", viruses) ~ "Tacheng tick virus XXX",
#                                grepl("Heartland virus", viruses) ~ "Heartland virus",
#                                grepl("Changping Tick Virus", viruses) ~ "Changping tick virus XXX",
#                                grepl("Bhanja virus", viruses) ~ "Bhanja virus XXX",
#                                grepl("Kupe virus", viruses) ~ "Kupe virus XXX",
#                                grepl("Lihan tick virus", viruses) ~ "Lihan tick virus XXX",
#                                grepl("Bourbon virus", viruses) ~ "Bourbon virus",
#                                grepl("Mukawa virus", viruses) ~ "Mukawa virus XXX",
#                                grepl("Punta Salinas", viruses) ~ "Punta Salinas XXX",
#                                grepl("Ohshima virus", viruses) ~ "Ohshima virus XXX",
#                                grepl("Hazara virus ", viruses) ~ "Hazara virus  XXX",
#                                grepl("Zaliv Terpenia virus", viruses) ~ "Zaliv Terpenia virus XXX",
#                                grepl("Paramushir virus", viruses) ~ "Paramushir virus XXX",
#                                grepl("Lesvos virus", viruses) ~ "Lesvos virus XXX",
#                                grepl("Kaisodi virus", viruses) ~ "Kaisodi virus XXX",
#                                grepl("Taggert virus", viruses) ~ "Taggert virus XXX",
#                                grepl("Bogoria virus", viruses) ~ "Bogoria virus XXX",
#                                grepl("Yongjia Tick Virus 2", viruses) ~ "Yongjia tick virus 2 XXX",
#                                grepl("Soldado virus", viruses) ~ "Soldado virus XXX",
#                                grepl("Kasokero virus", viruses) ~ "Kasokero virus XXX",
#                                grepl("Huaiyangshan virus", viruses) ~ "Huaiyangshan virus XXX",
#                                grepl("Powassan virus", viruses) ~ "Powassan virus",
#                                grepl("Tick-borne encephalitis virus", viruses) ~ "Tick-borne #encephalitis virus",
#                                grepl("Jingmen tick virus", viruses) ~ "Jingmen tick virus XXX",
#                                grepl("Karshi virus", viruses) ~ "Karshi virus XXX",
#                                grepl("Alkhumra hemorrhagic fever virus", viruses) ~ "Alkhumra #hemorrhagic fever virus XXX",
#                                grepl("Kyasanur Forest disease virus", viruses) ~ "Kyasanur Forest #disease virus  XXX",
#                                grepl("Komandory virus", viruses) ~ "Komandory virus XXX",
#                                grepl("Tjuloc virus", viruses) ~ "Tjuloc virus XXX",
#                                grepl("Rukutama virus", viruses) ~ "Rukutama virus XXX",
#                                grepl("Tillamook virus", viruses) ~ "Tillamook virus XXX",
#                                grepl("Rift Valley fever virus", viruses) ~ "Rift Valley fever virus #XXX",
#                                grepl("South Bay virus", viruses) ~ "South Bay virus XXX",
#                                grepl("Harz mountain virus", viruses) ~ "Harz mountain virus XXX",
#                                grepl("Guangxi tick virus", viruses) ~ "Guangxi tick virus XXX",
#                                grepl("West Nile virus", viruses) ~ "West Nile virus XXX",
#                                grepl("Alongshan virus", viruses) ~ "Alongshan virus XXX",
#                                grepl("Omsk hemorrhagic fever virus", viruses) ~ "Omsk hemorrhagic #fever virus XXX",
#                                grepl("Heilongjiang tick virus", viruses) ~ "Heilongjiang tick virus #XXX",
#                                grepl("Japanese encephalitis virus", viruses) ~ "Japanese #encephalitis virus XXX",
#                                grepl("Takachi virus", viruses) ~ "Takachi virus XXX",
#                                grepl("Louping ill virus", viruses) ~ "Louping ill virus XXX",
#                                grepl("Kindia tick virus", viruses) ~ "Kindia tick virus XXX",
#                                grepl("Deer tick virus", viruses) ~ "Deer tick virus XXX", # Powassan
#                                grepl("Mogiana tick virus", viruses) ~ "Mogiana tick virus XXX",
#                                grepl("Jingmenvirus", viruses) ~ "Jingmen tick virus XXX",
#                                grepl("Yamaguchi virus", viruses) ~ "Yamaguchi virus XXX",
#                                grepl("Yanggou tick virus", viruses) ~ "Yanggou tick virus XXX",
#                                grepl("Haseki tick virus", viruses) ~ "Haseki tick virus XXX",
#                                grepl("Issyk-Kul virus", viruses) ~ "Issyk-Kul virus XXX",
#                                grepl("Pustyn virus", viruses) ~ "Pustyn virus XXX",
#                                grepl("Tamdy virus", viruses) ~ "Tamdy virus XXX",
#                                grepl("Khasan virus", viruses) ~ "Khasan virus XXX",
#                                grepl("Batken virus", viruses) ~ "Batken virus XXX",
#                                grepl("Meihua Mountain virus", viruses) ~ "Meihua Mountain virus #XXX",
#                                grepl("Uumaja virus", viruses) ~ "Uumaja virus XXX",
#                                grepl("Nanning Phenu tick virus", viruses) ~ "Nanning Phenu tick #virus XXX",
#                                grepl("Yushu tick virus", viruses) ~ "Yushu tick virus XXX",
#                                grepl("Wenzhou Tick Virus", viruses) ~ "Wenzhou tick virus XXX",
#                                grepl("Langat virus", viruses) ~ "Langat virus XXX",
#                                grepl("Saumarez Reef virus", viruses) ~ "Saumarez Reef virus XXX",
#                                grepl("Finch creek virus", viruses) ~ "Finch creek virus XXX",
#                                grepl("Sapphire II virus", viruses) ~ "Sapphire II virus XXX",
#                                grepl("Zirqa virus", viruses) ~ "Zirqa virus XXX",
#                                grepl("Camel tick virus", viruses) ~ "Camel tick virus XXX",
#                                grepl("Ingleside virus", viruses) ~ "Ingleside virus XXX",
#                                grepl("North shore virus", viruses) ~ "North shore virus XXX",
#                                grepl("Shelly headland virus", viruses) ~ "Shelly headland virus #XXX",
#                                grepl("Avalon virus", viruses) ~ "Avalon virus XXX",
#                                grepl("Fuyun tick virus", viruses) ~ "Fuyun tick virus XXX",
#                                grepl("Dudero virus", viruses) ~ "Dudero virus XXX",
#                                grepl("Gravol virus", viruses) ~ "Gravol virus XXX",
#                                grepl("Karukera tick virus", viruses) ~ "Karukera tick virus XXX",
#                                grepl("North shore virus", viruses) ~ "North shore virus XXX",
#                                grepl("Reoviridae sp.", viruses) ~ "Reoviridae sp. YYY", # YYY for sp #not identified but want to get rid of na
#                                grepl("Ixovirus sp.",viruses) ~ "Ixovirus sp. YYY", 
#                                grepl("Peribunyaviridae sp.", viruses) ~ "Peribunyaviridae sp. YYY", 
#                                grepl("Rhabdoviridae sp.", viruses) ~ "Rhabdoviridae sp. YYY", 
#                                grepl("Orthomyxoviridae sp.", viruses) ~ "Orthomyxoviridae sp. YYY", 
#                                grepl("Nairoviridae sp.", viruses) ~ "Nairoviridae sp. YYY", 
#                                grepl("Phenuiviridae sp.", viruses) ~ "Phenuiviridae sp. YYY",
#                                grepl("Palyam virus", viruses) ~ "Palyam virus XXX",
#                                grepl("Great Island virus", viruses) ~ "Great Island virus XXX",
#                                grepl("Muko virus", viruses) ~ "Muko virus XXX",
#                                grepl("Aniva virus", viruses) ~ "Aniva virus XXX",
#                                grepl("Baku virus", viruses) ~ "Baku virus XXX",
#                                grepl("Matucare virus", viruses) ~ "Matucare virus XXX",
#                                grepl("Kundal virus", viruses) ~ "Kundal virus XXX",
#                                grepl("Nugget virus", viruses) ~ "Nugget virus XXX",
#                                grepl("Sandy Bay Virus", viruses) ~ "Sandy Bay virus XXX",
#                                grepl("Chobar Gorge virus", viruses) ~ "Chobar Gorge virus XXX",
#                                grepl("Gierle tick virus", viruses) ~ "Gierle tick virus XXX",
#                                grepl("Zoersel tick virus", viruses) ~ "Zoersel tick virus XXX",
#                                grepl("Lishui pangolin virus", viruses) ~ "Lishui pangolin virus #XXX",
#                                grepl("Tribec virus", viruses) ~ "Tribec virus XXX",
#                                grepl("Aleutian mink disease virus", viruses) ~ "Aleutian mink #disease virus XXX",
#                                grepl("New Minto virus", viruses) ~ "New Minto virus XXX",
#                                grepl("Imakoba tick virus", viruses) ~ "Imakoba tick virus XXX",
#                                grepl("Kabuto mountain virus", viruses) ~ "Kabuto mountain virus #XXX",
#                                grepl("Gissar virus", viruses) ~ "Gissar virus XXX",
#                                grepl("New Minto virus", viruses) ~ "New Minto virus XXX",
#                                grepl("Burana virus", viruses) ~ "Burana virus XXX",
#                                grepl("Bronnoya virus", viruses) ~ "Bronnoya virus XXX",
#                                grepl("Connecticut virus", viruses) ~ "Connecticut virus XXX",
#                                grepl("Grand Arbaud virus", viruses) ~ "Grand Arbaud virus XXX",
#                                grepl("Rondonia orthonairovirus", viruses) ~ "Rondonia #orthonairovirus YYY",
#                                grepl("Farallon virus", viruses) ~ "Farallon virus XXX",
#                                grepl("Silverwater virus", viruses) ~ "Silverwater virus XXX",
#                                grepl("Quaranjavirus sp.", viruses) ~ "Quaranjavirus sp. YYY",
#                                grepl("Precarious point virus", viruses) ~ "Precarious point virus #XXX",
#                                grepl("Vinegar Hill virus", viruses) ~ "Vinegar Hill virus XXX",
#                                grepl("Hunter Island virus", viruses) ~ "Hunter Island virus XXX",
#                                grepl("Kuriyama virus", viruses) ~ "Kuriyama virus XXX",
#                                grepl("Grotenhout virus", viruses) ~ "Grotenhout virus XXX",
#                                grepl("Gomselga virus", viruses) ~ "Gomselga virus XXX",
#                                grepl("Ngari virus", viruses) ~ "Ngari virus XXX",
#                                grepl("Wanowrie virus", viruses) ~ "Wanowrie virus XXX",
#                                grepl("Toyo virus", viruses) ~ "Toyo virus XXX",
#                                grepl("Ngari virus", viruses) ~ "Ngari virus XXX",
#                                grepl("Estero Real virus", viruses) ~ "Estero Real virus XXX",
#                                grepl("Iftin tick virus", viruses) ~ "Iftin tick virus XXX",
#                                grepl("Geran virus", viruses) ~ "Geran virus XXX",
#                                grepl("Zambezi tick virus", viruses) ~ "Zambezi tick virus XXX",
#                                grepl("Umea virus", viruses) ~ "Umea virus XXX",
#                                grepl("Bonden virus", viruses) ~ "Bonden virus XXX",
#                                grepl("Ubmeje virus", viruses) ~ "Ubmeje virus XXX",
#                                grepl("Manawa virus", viruses) ~ "Manawa virus XXX",
#                                grepl("Uukuniemi virus", viruses) ~ "Uukuniemi virus XXX",
#                                grepl("Palma virus", viruses) ~ "Palma virus XXX",
#                                grepl("Cencurut virus", viruses) ~ "Cencurut virus XXX",
#                                grepl("Sunday Canyon virus", viruses) ~ "Sunday Canyon virus XXX",
#                                grepl("Saphire II virus", viruses) ~ "Saphire II virus XXX",
#                                grepl("New Kent County virus", viruses) ~ "New Kent County virus #XXX",
#                                grepl("Laurel Lake virus", viruses) ~ "Laurel Lake virus XXX",
#                                grepl("Uukuniemi virus", viruses) ~ "Uukuniemi virus XXX",
#                                grepl("Stavropol virus", viruses) ~ "Stavropol virus XXX",
#                                grepl("Bronnoya virus", viruses) ~ "Bronnoya virus XXX",
#                                grepl("Dandong tick virus", viruses) ~ "Dandong tick virus XXX",
#                                grepl("Guyuan tick virus", viruses) ~ "Guyuan tick virus XXX",
#                                grepl("Heihe tick virus", viruses) ~ "Heihe tick virus XXX",
#                                grepl("Nanchang tick virus", viruses) ~ "Nanchang tick virus XXX",
#                                grepl("Tonghua tick virus", viruses) ~ "Tonghua tick virus XXX",
#                                grepl("Zhangye tick virus", viruses) ~ "Zhangye tick virus XXX",
#                                grepl("Balambala tick virus", viruses) ~ "Balambala tick virus XXX",
#                                grepl("Gakugsa tick virus", viruses) ~ "Gakugsa tick virus XXX",
#                                grepl("Lone Star virus", viruses) ~ "Lone Star virus XXX",
#                                grepl("Kuriyama virus", viruses) ~ "Kuriyama virus XXX",
#                                grepl("Kumlinge virus", viruses) ~ "Kumlinge virus XXX",
#                                grepl("Middelburg virus", viruses) ~ "Middelburg virus XXX",
#                                grepl("Meaban virus", viruses) ~ "Meaban virus XXX",
#                                grepl("Gadgets Gully virus", viruses) ~ "Gadgets Gully virus XXX",
#                                grepl("Royal Farm virus", viruses) ~ "Royal Farm virus XXX",
#                                grepl("Ngoye virus", viruses) ~ "Ngoye virus XXX",
#                                grepl("Jiutai virus", viruses) ~ "Jiutai virus XXX",
#                                grepl("Tyuleniy virus", viruses) ~ "Tyuleniy virus XXX",
#                                grepl("Kama virus", viruses) ~ "Kama virus XXX",
#                                grepl("Tyuleniy virus", viruses) ~ "Tyuleniy virus XXX",
#                                grepl("Tacheng tick virus", viruses) ~ "Tacheng tick virus XXX",
#                                grepl("Sikhote-Alin virus", viruses) ~ "Sikhote-Alin virus XXX",
#                                grepl("Trinbago virus", viruses) ~ "Trinbago virus XXX",
#                                grepl("Tacheng tick virus", viruses) ~ "Tacheng tick virus XXX",
#                                grepl("Flaviviridae sp.", viruses) ~ "Flaviviridae sp. YYY",
#                                grepl("Picornaviridae sp.", viruses) ~ "Picornaviridae sp. YYY",
#                                grepl("Turkish sheep encephalitis virus", viruses) ~ "Turkish sheep #encephalitis virus XXX",
#                                grepl("Greek goat encephalitis virus", viruses) ~ "Greek goat #encephalitis virus XXX",
#                                grepl("Hubei tick virus", viruses) ~ "Hubei tick virus XXX",
#                                grepl("Fennes virus", viruses) ~ "Fennes virus XXX",
#                                grepl("Bulatov virus", viruses) ~ "Bulatov virus XXX",
#                                grepl("Messner virus", viruses) ~ "Messner virus XXX",
#                                grepl("Gerbovich virus", viruses) ~ "Gerbovich virus XXX",
#                                grepl("Piguzov virus", viruses) ~ "Piguzov virus XXX",
#                                grepl("Vovk virus", viruses) ~ "Vovk virus XXX",
#                                grepl("Mackintosh virus", viruses) ~ "Mackintosh virus XXX",
#                                grepl("Ed virus", viruses) ~ "Ed virus XXX",
#                                grepl("Snesslinge virus", viruses) ~ "Snesslinge virus XXX",
#                                grepl("Agalma virus", viruses) ~ "Agalma virus XXX",
#                                grepl("Bolka virus", viruses) ~ "Bolka virus XXX",
#                                grepl("Qualyub virus", viruses) ~ "Qualyub virus XXX",
#                                grepl("Graso virus", viruses) ~ "Graso virus XXX",
#                                grepl("Hepelivirales sp.", viruses) ~ "Hepelivirales sp. YYY",
#                                grepl("Henan tick virus", viruses) ~ "Henan tick virus XXX",
#                                grepl("Valo virus", viruses) ~ "Valo virus XXX",
#                                grepl("Ganjam virus", viruses) ~ "Ganjam virus XXX",
#                                grepl("Timbillica virus", viruses) ~ "Timbillica virus XXX",
#                                grepl("Genoa virus", viruses) ~ "Genoa virus XXX",
#                                grepl("Liman tick virus", viruses) ~ "Liman tick virus XXX",
#                                 grepl("Canne point virus", viruses) ~ "Canne point virus XXX",
#                                 grepl("Collins beach virus", viruses) ~ "Collins beach virus XXX",
#                                 grepl("Fairfax lookout virus", viruses) ~ "Fairfax lookout virus #XXX",
#                                 grepl("Fairlight virus", viruses) ~ "Fairlight virus XXX",
#                                grepl("Jump rock virus", viruses) ~ "Jump rock virus XXX",
#                                 grepl("Manly virus", viruses) ~ "Manly virus XXX",
#                                 grepl("Nadgee virus", viruses) ~ "Nadgee virus XXX",
#                                 grepl("Old quarry swamp virus", viruses) ~ "Old quarry swamp virus #XXX",
#                                 grepl("Quarantine head virus", viruses) ~ "Quarantine head virus #XXX",
#                                grepl("Store beach virus", viruses) ~ "Store beach virus XXX",
#                                 grepl("Store beach virus", viruses) ~ "Store beach virus XXX",
#                                 grepl("Yambulla virus", viruses) ~ "Yambulla virus XXX",
#                                 grepl("Ronne virus", viruses) ~ "Ronne virus XXX",
#                                grepl("Fennes virus", viruses) ~ "Fennes virus XXX",
#                                 grepl("Gerbovich virus", viruses) ~ "Gerbovich virus XXX",
#                                 grepl("Bulatov virus", viruses) ~ "Bulatov virus XXX",
#                                 grepl("Vovk virus", viruses) ~ "Vovk virus XXX",
#                                grepl("Torhout tick virus", viruses) ~ "Torhout tick virus XXX",
#                                grepl("Great Saltee virus", viruses) ~ "Great Saltee virus XXX",
#                                grepl("Shanxi tick virus", viruses) ~ "Shanxi tick virus XXX",
#                                grepl("Hubei tick virus", viruses) ~ "Hubei tick virus XXX",
#                                grepl("Avalon Bres virus", viruses) ~ "Avalon Bres virus XXX",
#                                grepl("Henan tick virus", viruses) ~ "Henan tick virus XXX",
#                                grepl("Fort Crockett virus", viruses) ~ "Fort Crockett virus XXX"
#                                
#         )) 
#
#
## see if i can fix stragglers
#unique(tick_virus_all_cleanvirus$fix_viruses)
#tick_virus_all_cleanvirus[is.na(tick_virus_all_cleanvirus$fix_viruses),] # 553
#View(tick_virus_all_cleanvirus[is.na(tick_virus_all_cleanvirus$fix_viruses),]) 
#
### okay let's see what happens when i get rid of generic species
##tick_virus_all_cleanvirus %>% 
##  filter(grepl("Reoviridae sp.", viruses) |
#
#tick_virus_all_cleanvirus %>% 
#  filter(grepl("associated", viruses)) # 129 associated
#
#
#tick_virus_all %>% 
#  filter(references == "Unpublished") %>% 
#  tally() # 2933	
#
#tick_virus_all %>% 
#  filter(sample_type == "Culture") %>% 
#  tally() # 1395	

```

#Step 5. Save csv for all TBVs
(this is all tick-borne viruses - can be used for other projects)
```{r}
#write.csv(tick_virus_all_cleanvirus, "tickviruses/clean/zover_allviruses_20240921.csv")
```

#Step 6. Wrangle data for select TBVs

###Figure out which are the top reported TBVs on ZOVER
```{r}
# re-upload cleaned csv of all TBVs
#zover_all <- read.csv("tickviruses/clean/zover_allviruses_20240921.csv")
#
### first identify the current 30
#zover_all %>% 
#  group_by(viral_family,fix_viruses) %>% 
#  tally() %>% 
#  arrange(desc(n)) %>% 
#  # omit names that are meaningless
#  filter(fix_viruses != "NA" &
#           fix_viruses != "Rhabdoviridae sp. YYY" &
#           viral_family != "N/A" &
#           fix_viruses != "Flaviviridae sp. YYY" &
#           fix_viruses != "Nairoviridae sp. YYY" &
#           fix_viruses != "Peribunyaviridae sp. YYY" &
#           fix_viruses != "Orthomyxoviridae sp. YYY" &
#           fix_viruses != "Phenuiviridae sp. YYY" &
#           fix_viruses != "Bocaparvovirus" &
#           fix_viruses != "Phlebovirus XXX" &
#           n >30) 

```

###Fix names of the top reported TBVs
```{r}
## fix names of top ~20 viruses
#zover_all <- zover_all %>% 
#  mutate(fix_viruses = case_when(fix_viruses == "Deer tick virus XXX" ~ "Deer tick virus",
#                                 fix_viruses == "Crimean-Congo hemorrhagic fever"~ "Crimean-Congo #Hemorrhagic fever",
#                                 fix_viruses == "Severe fever with thrombocytopenia virus"~ "Severe #fever with thrombocytopenia syndrome",
#                                 fix_viruses == "Jingmen tick virus XXX"~ "Jingmen tick virus",
#                                 fix_viruses == "African swine fever"~ "African swine fever virus",
#                                 fix_viruses == "Dabieshan Tick"~ "Dabieshan tick virus",
#                                 fix_viruses == "Wuhan tick virus"~ "Wuhan tick virus",
#                                 fix_viruses == "Alongshan virus XXX"~ "Alongshan virus",
#                                 fix_viruses == "Lihan tick virus XXX"~ "Lihan tick virus",
#                                 fix_viruses == "Yanggou tick virus XXX"~ "Yanggou tick virus",
#                                 fix_viruses == "Mukawa virus XXX"~ "Mukawa virus",
#                                 fix_viruses == "Bole tick XXX"~ "Bole tick virus",
#                                 fix_viruses == "Haseki tick virus XXX"~ "Haseki tick virus",
#                                 fix_viruses == "Taishun Tick virus XXX"~ "Taishun tick virus",
#                                 fix_viruses == "Kyasanur Forest disease virus  XXX"~ "Kyasanur #Forest disease virus",
#                                 fix_viruses == "Guangxi tick virus XXX"~ "Guangxi tick virus",
#                                 fix_viruses == "Dugbe orthonairovirus XXX"~ "Dugbe orthonairovirus",
#                                 fix_viruses == "Huaiyangshan virus XXX"~ "Huaiyangshan virus",
#                                 fix_viruses == "Huangpi tick Virus"~ "Huangpi tick virus", 
#                                 fix_viruses == "Alkhumra hemorrhagic fever virus XXX" ~ "Alkhumra #hemorrhagic fever virus",
#                                 fix_viruses == "Bhanja virus XXX" ~ "Bhanja virus",
#                                 fix_viruses == "Louping ill virus XXX" ~ "Louping ill virus",
#                                 fix_viruses == "Omsk hemorrhagic fever virus XXX" ~ "Omsk #hemorrhagic fever virus",
#                                 TRUE ~ fix_viruses))  

#zover_all %>% filter(grepl("Saw", fix_viruses)) # helps find what is reported in column

```

###Filter to top ~15 viruses between 1990-2023
(this is scope of paper)

Top list based on Mansfield 2017, Lippi 2022, de la Fuente 2023
+ African swine fever
+ Alkhumra hemorrhagic fever virus
+ Bhanja virus
+ Bourbon virus
+ Colorado tick fever
+ Crimean-Congo hemorrhagic fever
+ Deer tick virus
+ Heartland virus
+ Jingmen tick virus
+ Kyasanur Forest disease virus
+ Louping ill virus
+ Lumpy skin disease
+ Nairobi sheep disease virus
+ Omsk hemorrhagic fever virus
+ Powassan virus
+ Sawgrass virus
+ Severe fever with thrombocytopenia virus
+ Tick-borne encephalitis virus

```{r}
#zover_toprecognized <- zover_all %>% # 7419 to 4213
#  filter(grepl("Heart",fix_viruses) |
#          grepl("Bourbon",fix_viruses) |
#          grepl("Severe",fix_viruses) |
#          grepl("Crimean",fix_viruses) |
#          grepl("Powassan",fix_viruses) |
#          grepl("Deer tick",fix_viruses) |
#          grepl("Kyasanur",fix_viruses) |
#          grepl("Alkhumra",fix_viruses) |
#          grepl("African swine",fix_viruses) |
#          grepl("Colorado",fix_viruses) |
#          grepl("Louping",fix_viruses) |
#          grepl("Nairobi",fix_viruses) |
#          grepl("Omsk",fix_viruses) |
#          grepl("Saw",fix_viruses) |
#          grepl("Bhanja",fix_viruses) |
#          grepl("Lumpy",fix_viruses) |
#          grepl("Tick-borne enceph", fix_viruses) |
#          grepl("Jingmen",fix_viruses)) %>% 
#  ### THINK ABOUT YEAR CUTOFF -- CHANGE HERE IF MAKE DIFFERENT DECISION ###
#  filter(collection_year > 1989) 
## got rid of Ganjam
#
#zover_toprecognized %>% 
#  group_by(sample_type) %>% 
#  tally()

```

#Step 7. Make model outcomes
Aggregate counts per country (tbv_total) & assign 1/0 per country (tbv_presence)
```{r}
#unique(zover_all$sampling_country)
#
### aggregate and assign binary values
#zover_toprecognized_countryaggregate <- zover_toprecognized %>% 
#  group_by(sampling_country) %>% 
#  tally() %>% 
#  # make sure country names match rnaturalearth
#  rename(admin = sampling_country) %>% 
#  right_join(worldmap, by = "admin") %>% 
#  mutate(iso_a3 = case_when(iso_a3 == "-99" ~ admin, # these are a handful of places like Naval Base #Guantanamo but want to make sure they get merged on to final product so will keep their admin name
#                            TRUE ~ iso_a3)) %>% 
#  # assign binomial count
#  mutate(tbv_presence = case_when(n > 0 ~ 1,
#                                  is.na(n) ~ 0),
#         n = case_when(is.na(n) ~ 0,
#                       TRUE ~ n)) %>% 
#  rename(tbv_total = n) %>% 
#  dplyr::select(admin, iso_a3, tbv_presence, tbv_total) 

```


#Step 8. Merge metadata & model outcomes
```{r}
# first slim the metadata to just relevant viruses
#zover_toprecognized_slim <- zover_toprecognized %>% 
#  dplyr::select(sampling_country, fix_viruses, # most important info
#                viral_family, type, # good info
#                continent,year_agg,  # for visualizations
#                from_tick, tick_genus, # for networks
#                references, sample_type,viruses, iso3c) %>% # info may be helpful for filtering
#  rename(admin = sampling_country)
#
### Merge aggregated country counts with relevant ZOVER metadata
#zover_toprecognized_complete <- zover_toprecognized_countryaggregate %>% 
#  right_join(zover_toprecognized_slim, by = "admin") # keep all rows in zover_toprecog_zlim
```

#Step 9. Add citation counts using easyPubMed
(probably should have created a function but did brute force - improve upon)
```{r}
## check which version gets more hits
get_pubmed_ids("Tick-borne encephalitis")$Count
# African swine 4125 / African swine fever 3474
# Alkhumra hemorrhagic 34 / Alkhumra hemorrhagic fever 32
# Bourbon 724 / Bourbon virus 43
# Colorado tick 581 / Colorado tick fever 288
# Crimean-congo hemorrhagic 2131 / Crimean-congo hemorrhagic fever 2119
# Deer tick 10316 / Deer tick virus 1590
# Heartland 3364 / Heartland virus 111
# Jingmen tick 56 / Jingmen tick virus 56
# Kyasanur forest 298 / Kyasanur forest disease 297
# Lumpy skin 711 / Lumpy skin disease 681
# Nairobi sheep 399 / Nairobi sheep disease 226
# Powassan 411 / Powassan virus 4211
# Severe fever with thrombocytopenia syndrome 1983 / Severe fever with thrombocytopenia syndrome virus 1367
# Tick-borne encephalitis 6561 / Tick-borne encephalitis virus 4954 

## pull list of names
#toprecognized_list <- zover_toprecognized_complete %>% 
#  group_by(fix_viruses) %>% 
#  slice(1) %>% 
#  dplyr::select(fix_viruses) %>% 
#  # fix names to match NCBI Taxonomy Browser
#  # https://www.ncbi.nlm.nih.gov/Taxonomy/Browser/wwwtax.cgi
#  mutate(fix_viruses = case_when(fix_viruses == "African swine fever virus" ~ "African swine",
#                                 fix_viruses == "Alkhumra hemorrhagic fever virus" ~ "Alkhumra #hemorrhagic",
#                                 fix_viruses == "Colorado tick fever" ~ "Colorado tick fever",
#                                 fix_viruses == "Crimean-Congo Hemorrhagic fever" ~ "Crimean-Congo #hemorrhagic fever",
#                                 fix_viruses == "Deer tick virus" ~ "Deer tick virus",
#                                 fix_viruses == "Kyasanur Forest disease virus" ~ "Kyasanur forest #disease",
#                                 fix_viruses == "Lumpy skin disease" ~ "Lumpy skin disease",
#                                 fix_viruses == "Nairobi sheep disease virus" ~ "Nairobi sheep #disease",
#                                 fix_viruses == "Severe fever with thrombocytopenia syndrome" ~ #"Severe fever with thrombocytopenia syndrome",
#                                 fix_viruses == "Tick-borne encephalitis virus" ~ "Tick-borne #encephalitis",
#                                 TRUE ~ fix_viruses))


## collect citations per virus species
# one way of doing this
# cites = c()
# for(i in 1:length(toprecognized_list$fix_viruses)) {
#   counts = as.numeric(as.character(get_pubmed_ids(gsub("_","-", toprecognized_list$fix_viruses[i]))$Count))
#   cites[i] = counts
#   print(paste(i,"/", nrow(toprecognized_list)))
# }

# looks for variations in virus or viral
#vcites = c()
#for(i in 1:length(toprecognized_list$fix_viruses)) {
#  x = gsub("_","-", toprecognized_list$fix_viruses[i])
#  x = paste("(",x,")",sep="")
#  x = paste(x,"AND (virus OR viral)")
#  counts = as.numeric(as.character(get_pubmed_ids(x)$Count))
#  vcites[i]=counts
#  print(paste(i,"/", nrow(toprecognized_list)))
#}
#
#
## turn into dataframe
## virus citations
#vcites_df <- as.data.frame(vcites)
## bind names with virus citations
#toprecognized_pubmedcitations <- cbind(toprecognized_list, vcites_df)

```

#Step 10. Aggregate citation count per country
```{r}
## aggregate
#zover_toprecognized_countryaggregate_pubmedcitations <- 
#zover_toprecognized_complete %>% 
#  mutate(fix_viruses = case_when(fix_viruses == "African swine fever virus" ~ "African swine",
#                                 fix_viruses == "Alkhumra hemorrhagic fever virus" ~ "Alkhumra #hemorrhagic",
#                                 fix_viruses == "Colorado tick fever" ~ "Colorado tick fever",
#                                 fix_viruses == "Crimean-Congo Hemorrhagic fever" ~ "Crimean-Congo #hemorrhagic fever",
#                                 fix_viruses == "Deer tick virus" ~ "Deer tick virus",
#                                 fix_viruses == "Kyasanur Forest disease virus" ~ "Kyasanur forest #disease",
#                                 fix_viruses == "Lumpy skin disease" ~ "Lumpy skin disease",
#                                 fix_viruses == "Nairobi sheep disease virus" ~ "Nairobi sheep #disease",
#                                 fix_viruses == "Severe fever with thrombocytopenia syndrome" ~ #"Severe fever with thrombocytopenia syndrome",
#                                 fix_viruses == "Tick-borne encephalitis virus" ~ "Tick-borne #encephalitis",
#                                 TRUE ~ fix_viruses)) %>% 
#left_join(toprecognized_pubmedcitations, by = "fix_viruses") %>% 
#  group_by(admin) %>% 
#  summarise(pubmedcitations_percounty_total = sum(vcites))
#
### add citation weight to each country
## Merge pubmed citation list with complete zover top recognized list
#zover_toprecognized_complete_pubmed <- zover_toprecognized_complete %>% 
#  left_join(zover_toprecognized_countryaggregate_pubmedcitations, by = "admin")
#
## Merge pubmed citation list with slim zover top recognized list
#zover_toprecognized_slim_pubmed <- zover_toprecognized_countryaggregate %>% 
#  left_join(zover_toprecognized_countryaggregate_pubmedcitations, by = "admin")
```

#Step 11. Save csv for select TBVs
(this is for select tick-borne viruses used in paper)
```{r}
#write.csv(zover_toprecognized_slim_pubmed, file = "tickviruses/clean/zover_toprecognized_slim_pubmed_20251103.csv")
  
#write.csv(zover_toprecognized_complete_pubmed , file = "tickviruses/clean/zover_toprecognized_complete_pubmed _20240921.csv")
```

