---
title: "2_ZOVEROutcomes_RawtoClean"
output: html_document
date: "2024-09-20"
---

**Purpose** 
To transform raw data from ZOVER into somewhat cleaned csv to further filter into final outcome variables

**Steps**

1) Upload raw data from ZOVER
2) Quick coarse clean for EDA purposes
3) Add country metadata
4) More refined cleaning for collection year
5) More EDA post cleaning
6) Make changes to viral families names
7) Make changes to individual virus names
8) Write final csv of clean-ish tick-viruses


###Final CSV: tick_viruses_clean_20240921.csv
^^ this will be filtered to final outcomes for BRT analysis

*Note: The ZOVER data is freely available under the Creative Commons Attribution-NonCommercial (CC BY-NC) license version 4.0 for personal and public non-commercial, research or academic use by individuals at academic, government or non-profit institutions. Users intending to use ZOVER data for commercial purposes should contact us via the official email.*

+ full ZOVER website (http://www.mgc.ac.cn/cgi-bin/ZOVER/main.cgi)
+ tick-virus website (http://www.mgc.ac.cn/cgi-bin/ZOVER/mainTable.cgi?db=tick)

Once on tick-virus wesbite, click on subgroups of tick-viruses (e.g. ssDNA), click "show all", click "save this table", open in personal excel, save as csv to Rproj

This data was accessed 2024-09-21

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

## upload necessary packages
library(ggplot2)
library(tidyverse)
library(readr)
library(janitor)
library(lubridate)
library(stringr)
library(countrycode)
library(rnaturalearth)
```

##Step 1. Upload raw data
```{r}
## upload individual tables for Tick-Classified Viruses

dsDNA_noRNA <- read_csv("tickviruses/raw/Tick-associated viruses (dsDNA viruses, no RNA stage) .csv", skip = 1) %>% clean_names()
dsRNA <- read_csv("tickviruses/raw/Tick-associated viruses (dsRNA viruses) .csv", skip = 1) %>% clean_names()
ssDNA <- read_csv("tickviruses/raw/Tick-associated viruses (ssDNA viruses) .csv", skip = 1) %>% clean_names()
ssRNA_negstrands <- read_csv("tickviruses/raw/Tick-associated viruses (ssRNA negative-strand viruses) .csv", skip = 1) %>% clean_names()
ssRNA_posstrands <- read_csv("tickviruses/raw/Tick-associated viruses (ssRNA positive-strand viruses, no DNA stage) .csv", skip = 1) %>% clean_names()
unlcassified <- read_csv("tickviruses/raw/Tick-associated viruses (unclassified viruses) .csv", skip = 1) %>% clean_names()

## add unique identifier for each database
dsDNA_noRNA$type <- "dsDNA_noRNA"
dsRNA$type <- "dsRNA"
ssDNA$type <- "ssDNA"
ssRNA_negstrands$type <- "ssRNA_negstrands"
ssRNA_posstrands$type <- "ssRNA_posstrands"
unlcassified$type <- "unlcassified"

## combine all csv files into 1 csv file
tick_viruses_raw <- rbind(dsDNA_noRNA, dsRNA, ssDNA, ssRNA_negstrands, ssRNA_posstrands, unlcassified )

# add count column for EDA
tick_viruses_raw$count <- 1

# check structure of dataframe
str(tick_viruses_raw) # 8,110 x 11 (~500 more since accessed 2023-09-24)
```

##Step 2. Quick coarse clean for EDA
- column names, check levels, fix some pecularities about virus names but will need to do this more systematically
```{r}
# clean up column names and data types
tick_virus_all <- tick_viruses_raw %>% 
  clean_names() %>% 
  mutate_at(c('collection_year', 'sampling_country','type'),
            as.factor) 

# check that levels are consistent
length(unique(tick_virus_all$viruses)) # 7611
length(unique(tick_virus_all$viral_family)) # 25
length(unique(tick_virus_all$from_tick)) # 152
length(unique(tick_virus_all$tick_genus)) # 13

# split up viruses into more manageable analysis chunks
tick_virus_all[,12:13] <-str_split_fixed(tick_virus_all$viruses, c("virus|Virus|sp."), 2)
#tick_virus_all[,12:13] <- str_split_fixed(tick_virus_all$viruses, "virus", 2)
names(tick_virus_all)[12] <- "virus_abbrev"
names(tick_virus_all)[13] <- "virus_strain"

# check how much more this need works
# unique(tick_virus_all$virus_abbrev)

# trim extra white space at right side of label
tick_virus_all$virus_abbrev <- str_trim(tick_virus_all$virus_abbrev,"right")

## lump some virus abbreviations together
# CCHF
tick_virus_all$virus_abbrev[(tick_virus_all$virus_abbrev) =="Crimean-Congo hemorrhagic fever orthonairo"] <- "Crimean-Congo hemorrhagic fever"

# Tick Phlebovirus
tick_virus_all$virus_abbrev[(tick_virus_all$virus_abbrev) =="Phlebo"] <- "Tick phlebo"
tick_virus_all$virus_abbrev[(tick_virus_all$virus_abbrev) =="unclassified Phlebo"] <- "Tick phlebo"
tick_virus_all$virus_abbrev[(tick_virus_all$virus_abbrev) =="Jingmen"] <- "Jingmen tick"


```

##Step 3. Add spatial metadata
```{r}
# add iso3c code
tick_virus_all$iso3c <- countrycode(tick_virus_all$sampling_country, 
            origin = 'country.name',
            destination = 'iso3c')

# add continent 
 tick_virus_all <-  tick_virus_all %>%
   mutate(continent = countrycode(iso3c, origin = "iso3c",
                                  destination = "continent"))
 
 
# upload world map
world <- ne_countries(scale = "large", returnclass = "sf")

```

##Step 4. More refined cleaning

### Fix dates

The steps I decided to do:

1. observations with ~ = if it's same year fix to same year (ie 2019~2019 is just 2019)

2. observations with ~ = if it's across 2 years, go with the first (ie 2018-2019 is 2018)

3. see how many more observations with ~ = either omit or make a judgement call

4. separate years into every 5 years?

```{r}

unique(tick_virus_all$collection_year)

## step 1. if same year fix to just 1 year

tick_virus_all$collection_year[(tick_virus_all$collection_year) == "2022~2022"] <- "2022"

tick_virus_all$collection_year[(tick_virus_all$collection_year) == "2020~2020"] <- "2020"

tick_virus_all$collection_year[(tick_virus_all$collection_year) == "2019~2019"] <- "2019"

tick_virus_all$collection_year[(tick_virus_all$collection_year) == "2017~2017"] <- "2017"

tick_virus_all$collection_year[(tick_virus_all$collection_year) == "2016~2016"] <- "2016"

tick_virus_all$collection_year[(tick_virus_all$collection_year) == "2015~2015"] <- "2015"

tick_virus_all$collection_year[(tick_virus_all$collection_year) == "2013~2013"] <- "2013"

tick_virus_all$collection_year[(tick_virus_all$collection_year) == "2005~2005"] <- "2005"


## step 2. if only 2 year span choose first year
tick_virus_all$collection_year[(tick_virus_all$collection_year) == "2018~2019"] <- "2018"

tick_virus_all$collection_year[(tick_virus_all$collection_year) == "2016~2017"] <- "2016"

tick_virus_all$collection_year[(tick_virus_all$collection_year) == "2015~2016"] <- "2015"

tick_virus_all$collection_year[(tick_virus_all$collection_year) == "2014~2015"] <- "2014"

tick_virus_all$collection_year[(tick_virus_all$collection_year) == "2011~2012"] <- "2011"

tick_virus_all$collection_year[(tick_virus_all$collection_year) == "2004~2005"] <- "2004"

tick_virus_all$collection_year[(tick_virus_all$collection_year) == "2000~2001"] <- "2000"

tick_virus_all$collection_year[(tick_virus_all$collection_year) == "1997~1998"] <- "1997"

## step 3a. how many other 1+ year strings are there?
tick_virus_all %>% 
  filter(grepl("~", collection_year)) # 123 observations

## step 3b. If it's only 2 year difference choose the middle year

tick_virus_all$collection_year[(tick_virus_all$collection_year) == "2019~2021"] <- "2020"

tick_virus_all$collection_year[(tick_virus_all$collection_year) == "2013~2015"] <- "2014"

tick_virus_all$collection_year[(tick_virus_all$collection_year) == "2012~2014"] <- "2012"

tick_virus_all$collection_year[(tick_virus_all$collection_year) == "2009~2011"] <- "2010"





## now there's 77 observations with years that are reported 2+ years together



## step 4. edit other random erros
tick_virus_all$collection_year[(tick_virus_all$collection_year) == "0199"] <- "1990"

tick_virus_all %>% 
  filter(collection_year == "ND") # 462 years not reported


### okay let's try to map this out as best as possible and make a decision about the 77 observations with 2+ year aggregation reported

unique(tick_virus_all$collection_year)

tick_virus_all %>% 
  filter(grepl("~", collection_year)) # 56 observations

dim(tick_virus_all)# 8110 × 15
```

Let's try to aggregate observations by 5 years
```{r}
tick_virus_all <- tick_virus_all %>% 
  mutate(collection_year = as.numeric(as.character(collection_year))) %>% 
  filter(!(grepl("~", collection_year)) & collection_year != "ND") %>% 
  mutate(year_agg = case_when(
    collection_year >= 2020 ~ "2020",
    collection_year >= 2015 ~ "2015",
    collection_year >= 2010 ~ "2010",
    collection_year >= 2005 ~ "2005",
    collection_year >= 2000 ~ "2000",
    collection_year >= 1995 ~ "1995",
    collection_year >= 1990 ~ "1990",
    collection_year >= 1985 ~ "1985",
    collection_year >= 1980 ~ "1980",
    collection_year >= 1975 ~ "1975",
    collection_year >= 1970 ~ "1970",
    collection_year >= 1965 ~ "1965",
    collection_year >= 1960 ~ "1960",
    collection_year >= 1955 ~ "1955",
    collection_year >= 1950 ~ "1950",
    collection_year < 1950 ~ "pre-1950"))

tick_virus_all %>% 
  filter(year_agg == "pre-1950") # only 3 left so I feel fine leaving it as that
```

##Step 5. More EDA post cleaning

check all countries are cleaned up
```{r}
unique(tick_virus_all$sampling_country) # 94/195 total

# let's see what's ND
View(tick_virus_all %>%  # 28
  filter(sampling_country == "ND") %>% 
    group_by(references) %>% 
    tally()) # 28 observations not listed.. checking reference

# many of these ND are cultures or unpublished.. maybe should get rid of both

tick_virus_all %>% 
  filter(sample_type == "Culture") #1,433 observations are cultured.. probably should remove these but worry about that once I get to analysis
```

check all viral familiy are cleaned up
```{r}
# first check unique viral families
unique(tick_virus_all$viral_family) # 23 unique families + 1 NA


# figure out what's NA
tick_virus_all %>% 
  filter(viral_family == "N/A") %>%  # 750 observations ... after 4 fixes its 467
  group_by(sampling_country,viral_family) %>% 
  tally() %>% # okay so a lot come from China, let's see if there's a common virus I can convert to viral family
 arrange(desc(n))

# fixing NA from China
tick_virus_all %>% 
  filter(viral_family == "N/A" & sampling_country == "China") %>% 
  group_by(virus_abbrev) %>% 
  tally() %>%  # bole tick bole,  hyalomma asisticum, 
  arrange(desc(n))

tick_virus_all %>%
  filter(grepl("Bole", viruses)) %>% 
  group_by(viral_family) %>% 
  tally() %>% 
  arrange(desc(n))#  okay so bole viruses can be phenuiviridae or rhabdoviridae... 63 observations are NA why is that

tick_virus_all %>%
  filter(grepl("Bole", viruses) & viral_family == "N/A") 

tick_virus_all %>%
  filter(grepl("Jingmen", viruses)) # looks like it's flaviviridae

tick_virus_all %>%
  filter(grepl("Wuhan", viruses)) # looks like it's Rhabdoviridae

# fixing NA from Australia
tick_virus_all %>% 
  filter(viral_family == "N/A" & sampling_country == "Australia") %>% 
  group_by(virus_abbrev) %>% # so nothing in australia has viral family resolved
  tally() %>%  
  arrange(desc(n))

tick_virus_all %>%
  filter(grepl("Shelly headland", viruses)) 

# fixing NAs from Russia
tick_virus_all %>% 
  filter(viral_family == "N/A" & sampling_country == "Russia") %>% 
  group_by(virus_abbrev)


## okay fixed some let's check main NAs left 467.. will try to par this down further later
tick_virus_all %>% 
  filter(viral_family == "N/A") %>%  # & sampling_country != "Australia") %>% 
  group_by(virus_abbrev) %>% 
  tally() %>% 
  arrange(desc(n)) # bole tick 55

```

##Step 6. Make more changes to viral families

Direct editing viral family (go back to re-run this with code)
```{r}
# first fix
tick_virus_all <- 
tick_virus_all %>%
  mutate(viral_family = if_else(
    grepl("Jingmen", viruses, fixed = TRUE),
    "Flaviviridae", viral_family
  )) 
  # filter(grepl("Jingmen", viruses))

# second fix
tick_virus_all <- 
tick_virus_all %>%
  mutate(viral_family = if_else(
    grepl("Wuhan", viruses, fixed = TRUE),
    "Rhabdoviridae", viral_family
  )) #%>% 
  #filter(grepl("Wuhan", viruses))


# 3rd fix
tick_virus_all <- 
  tick_virus_all %>%
  mutate(viral_family = if_else(
    grepl("Haseki", viruses, fixed = TRUE),
    "Flaviviridae", viral_family
  ))  
  #filter(grepl("Haseki", viruses))


# 4rd fix
tick_virus_all <- 
tick_virus_all %>%
  mutate(viral_family = if_else(
    grepl("Yanggou", viruses, fixed = TRUE),
    "Flaviviridae", viral_family
  ))  
  #filter(grepl("Yanggou", viruses))
```


##Step 7. Make changes to individual virus names

Coarse way to do this but for the purposes of the proposed papers, this is fine
```{r}
# see all unique
unique(tick_virus_all$virus_abbrev)

# pull out the most common ones

# test this function first
#tick_virus_all %>%
#  mutate(fix_viruses = case_when(grepl("Kemerovo virus", viruses) ~ "Kemerovo tickborne viral fever")) %>% 
#  dplyr::select(c(2,18))


tick_virus_all_cleanvirus <- tick_virus_all %>%
  mutate(fix_viruses = case_when(grepl("African swine fever", viruses) ~ "African swine fever",
                                 grepl("Colorado tick fever", viruses) ~ "Colorado tick fever",
                                grepl("Lumpy skin disease", viruses) ~ "Lumpy skin disease",
                                grepl("Kemerovo virus", viruses) ~ "Kemerovo tickborne viral fever",
                                grepl("Pseudocowpox virus", viruses) ~ "Pseudocowpox virus",
                                grepl("Eyach virus", viruses) ~ "Eyach virus", #TBE?
                                grepl("Crimean-Congo hemorrhagic fever", viruses) ~ "Crimean-Congo hemorrhagic fever",
                                grepl("Okutama tick virus", viruses) ~ "Okutama tick virus",
                                grepl("bole tick", viruses) ~ "Bole tick XXX",
                                grepl("Bole Tick", viruses) ~ "Bole tick XXX",
                                grepl("Bole tick", viruses) ~ "Bole tick XXX",
                                grepl("Bocaparvovirus", viruses) ~ "Bocaparvovirus",
                                grepl("Severe fever with thrombocytopenia virus", viruses) ~ "Severe fever with thrombocytopenia virus",
                                grepl("Nairobi sheep disease virus", viruses) ~ "Nairobi sheep disease virus",
                                grepl("Dabieshan Tick", viruses) ~ "Dabieshan Tick",
                                grepl("St Croix River virus", viruses) ~ "St Croix River virus",
                                grepl("Sawgrass", viruses) ~ "Sawgrass virus",
                                grepl("Phlebovirus", viruses) ~ "Phlebovirus XXX",
                                grepl("Taishun Tick Virus", viruses) ~ "Taishun Tick virus XXX",
                                grepl("phlebovirus", viruses) ~ "Phlebovirus XXX",
                                grepl("Wuhan Tick Virus 1", viruses) ~ "Wuhan tick virus 1",
                                grepl("Wuhan Tick Virus 2", viruses) ~ "Wuhan tick virus 2",
                                grepl("Wuhan tick virus 2", viruses) ~ "Wuhan tick virus 2",
                                grepl("Tarumizu tick virus", viruses) ~ "Tarumizu tick virus",
                                grepl("Guangdong tick orbivirus", viruses) ~ "Guangdong tick orbivirus",
                                grepl("Xinjiang tick orbivirus", viruses) ~ "Xinjiang tick orbivirus",
                                grepl("Okhotskiy virus", viruses) ~ "Okhotskiy virus",
                                grepl("Huangpi Tick Virus", viruses) ~ "Huangpi tick Virus",
                                grepl("Dugbe orthonairovirus ", viruses) ~ "Dugbe orthonairovirus XXX",
                                grepl("Tacheng Tick Virus 1", viruses) ~ "Tacheng tick virus 1 XXX",
                                grepl("Yezo virus", viruses) ~ "Yezo virus XXX",
                                grepl("Zhangzhou tick virus 1", viruses) ~ "Zhangzhou tick virus 1 XXX",
                                grepl("Songling virus", viruses) ~ "Songling virus XXX",
                                grepl("Wad Medani virus", viruses) ~ "Wad Medani virus XXX",
                                grepl("Tacheng Tick Virus ", viruses) ~ "Tacheng Tick virus XXX",                                 
                                grepl("Odaw virus", viruses) ~ "Odaw virus XXX",
                                grepl("Guertu virus", viruses) ~ "Guertu virus XXX",
                                grepl("Shibuyunji virus", viruses) ~ "Shibuyunji virus XXX",
                                grepl("Quaranfil quaranjavirus", viruses) ~ "Quaranfil quaranjavirus XXX",
                                grepl("Tacheng Tick Virus", viruses) ~ "Tacheng tick virus XXX",
                                grepl("Heartland virus", viruses) ~ "Heartland virus",
                                grepl("Changping Tick Virus", viruses) ~ "Changping tick virus XXX",
                                grepl("Bhanja virus", viruses) ~ "Bhanja virus XXX",
                                grepl("Kupe virus", viruses) ~ "Kupe virus XXX",
                                grepl("Lihan tick virus", viruses) ~ "Lihan tick virus XXX",
                                grepl("Bourbon virus", viruses) ~ "Bourbon virus",
                                grepl("Mukawa virus", viruses) ~ "Mukawa virus XXX",
                                grepl("Punta Salinas", viruses) ~ "Punta Salinas XXX",
                                grepl("Ohshima virus", viruses) ~ "Ohshima virus XXX",
                                grepl("Hazara virus ", viruses) ~ "Hazara virus  XXX",
                                grepl("Zaliv Terpenia virus", viruses) ~ "Zaliv Terpenia virus XXX",
                                grepl("Paramushir virus", viruses) ~ "Paramushir virus XXX",
                                grepl("Lesvos virus", viruses) ~ "Lesvos virus XXX",
                                grepl("Kaisodi virus", viruses) ~ "Kaisodi virus XXX",
                                grepl("Taggert virus", viruses) ~ "Taggert virus XXX",
                                grepl("Bogoria virus", viruses) ~ "Bogoria virus XXX",
                                grepl("Yongjia Tick Virus 2", viruses) ~ "Yongjia tick virus 2 XXX",
                                grepl("Soldado virus", viruses) ~ "Soldado virus XXX",
                                grepl("Kasokero virus", viruses) ~ "Kasokero virus XXX",
                                grepl("Huaiyangshan virus", viruses) ~ "Huaiyangshan virus XXX",
                                grepl("Powassan virus", viruses) ~ "Powassan virus",
                                grepl("Tick-borne encephalitis virus", viruses) ~ "Tick-borne encephalitis virus",
                                grepl("Jingmen tick virus", viruses) ~ "Jingmen tick virus XXX",
                                grepl("Karshi virus", viruses) ~ "Karshi virus XXX",
                                grepl("Alkhumra hemorrhagic fever virus", viruses) ~ "Alkhumra hemorrhagic fever virus XXX",
                                grepl("Kyasanur Forest disease virus", viruses) ~ "Kyasanur Forest disease virus  XXX",
                                grepl("Komandory virus", viruses) ~ "Komandory virus XXX",
                                grepl("Tjuloc virus", viruses) ~ "Tjuloc virus XXX",
                                grepl("Rukutama virus", viruses) ~ "Rukutama virus XXX",
                                grepl("Tillamook virus", viruses) ~ "Tillamook virus XXX",
                                grepl("Rift Valley fever virus", viruses) ~ "Rift Valley fever virus XXX",
                                grepl("South Bay virus", viruses) ~ "South Bay virus XXX",
                                grepl("Harz mountain virus", viruses) ~ "Harz mountain virus XXX",
                                grepl("Guangxi tick virus", viruses) ~ "Guangxi tick virus XXX",
                                grepl("West Nile virus", viruses) ~ "West Nile virus XXX",
                                grepl("Alongshan virus", viruses) ~ "Alongshan virus XXX",
                                grepl("Omsk hemorrhagic fever virus", viruses) ~ "Omsk hemorrhagic fever virus XXX",
                                grepl("Heilongjiang tick virus", viruses) ~ "Heilongjiang tick virus XXX",
                                grepl("Japanese encephalitis virus", viruses) ~ "Japanese encephalitis virus XXX",
                                grepl("Takachi virus", viruses) ~ "Takachi virus XXX",
                                grepl("Louping ill virus", viruses) ~ "Louping ill virus XXX",
                                grepl("Kindia tick virus", viruses) ~ "Kindia tick virus XXX",
                                grepl("Deer tick virus", viruses) ~ "Deer tick virus XXX", # Powassan
                                grepl("Mogiana tick virus", viruses) ~ "Mogiana tick virus XXX",
                                grepl("Jingmenvirus", viruses) ~ "Jingmen tick virus XXX",
                                grepl("Yamaguchi virus", viruses) ~ "Yamaguchi virus XXX",
                                grepl("Yanggou tick virus", viruses) ~ "Yanggou tick virus XXX",
                                grepl("Haseki tick virus", viruses) ~ "Haseki tick virus XXX",
                                grepl("Issyk-Kul virus", viruses) ~ "Issyk-Kul virus XXX",
                                grepl("Pustyn virus", viruses) ~ "Pustyn virus XXX",
                                grepl("Tamdy virus", viruses) ~ "Tamdy virus XXX",
                                grepl("Khasan virus", viruses) ~ "Khasan virus XXX",
                                grepl("Batken virus", viruses) ~ "Batken virus XXX",
                                grepl("Meihua Mountain virus", viruses) ~ "Meihua Mountain virus XXX",
                                grepl("Uumaja virus", viruses) ~ "Uumaja virus XXX",
                                grepl("Nanning Phenu tick virus", viruses) ~ "Nanning Phenu tick virus XXX",
                                grepl("Yushu tick virus", viruses) ~ "Yushu tick virus XXX",
                                grepl("Wenzhou Tick Virus", viruses) ~ "Wenzhou tick virus XXX",
                                grepl("Langat virus", viruses) ~ "Langat virus XXX",
                                grepl("Saumarez Reef virus", viruses) ~ "Saumarez Reef virus XXX",
                                grepl("Finch creek virus", viruses) ~ "Finch creek virus XXX",
                                grepl("Sapphire II virus", viruses) ~ "Sapphire II virus XXX",
                                grepl("Zirqa virus", viruses) ~ "Zirqa virus XXX",
                                grepl("Camel tick virus", viruses) ~ "Camel tick virus XXX",
                                grepl("Ingleside virus", viruses) ~ "Ingleside virus XXX",
                                grepl("North shore virus", viruses) ~ "North shore virus XXX",
                                grepl("Shelly headland virus", viruses) ~ "Shelly headland virus XXX",
                                grepl("Avalon virus", viruses) ~ "Avalon virus XXX",
                                grepl("Fuyun tick virus", viruses) ~ "Fuyun tick virus XXX",
                                grepl("Dudero virus", viruses) ~ "Dudero virus XXX",
                                grepl("Gravol virus", viruses) ~ "Gravol virus XXX",
                                grepl("Karukera tick virus", viruses) ~ "Karukera tick virus XXX",
                                grepl("North shore virus", viruses) ~ "North shore virus XXX",
                                grepl("Reoviridae sp.", viruses) ~ "Reoviridae sp. YYY", # YYY for sp not identified but want to get rid of na
                                grepl("Ixovirus sp.",viruses) ~ "Ixovirus sp. YYY", 
                                grepl("Peribunyaviridae sp.", viruses) ~ "Peribunyaviridae sp. YYY", 
                                grepl("Rhabdoviridae sp.", viruses) ~ "Rhabdoviridae sp. YYY", 
                                grepl("Orthomyxoviridae sp.", viruses) ~ "Orthomyxoviridae sp. YYY", 
                                grepl("Nairoviridae sp.", viruses) ~ "Nairoviridae sp. YYY", 
                                grepl("Phenuiviridae sp.", viruses) ~ "Phenuiviridae sp. YYY",
                                grepl("Palyam virus", viruses) ~ "Palyam virus XXX",
                                grepl("Great Island virus", viruses) ~ "Great Island virus XXX",
                                grepl("Muko virus", viruses) ~ "Muko virus XXX",
                                grepl("Aniva virus", viruses) ~ "Aniva virus XXX",
                                grepl("Baku virus", viruses) ~ "Baku virus XXX",
                                grepl("Matucare virus", viruses) ~ "Matucare virus XXX",
                                grepl("Kundal virus", viruses) ~ "Kundal virus XXX",
                                grepl("Nugget virus", viruses) ~ "Nugget virus XXX",
                                grepl("Sandy Bay Virus", viruses) ~ "Sandy Bay virus XXX",
                                grepl("Chobar Gorge virus", viruses) ~ "Chobar Gorge virus XXX",
                                grepl("Gierle tick virus", viruses) ~ "Gierle tick virus XXX",
                                grepl("Zoersel tick virus", viruses) ~ "Zoersel tick virus XXX",
                                grepl("Lishui pangolin virus", viruses) ~ "Lishui pangolin virus XXX",
                                grepl("Tribec virus", viruses) ~ "Tribec virus XXX",
                                grepl("Aleutian mink disease virus", viruses) ~ "Aleutian mink disease virus XXX",
                                grepl("New Minto virus", viruses) ~ "New Minto virus XXX",
                                grepl("Imakoba tick virus", viruses) ~ "Imakoba tick virus XXX",
                                grepl("Kabuto mountain virus", viruses) ~ "Kabuto mountain virus XXX",
                                grepl("Gissar virus", viruses) ~ "Gissar virus XXX",
                                grepl("New Minto virus", viruses) ~ "New Minto virus XXX",
                                grepl("Burana virus", viruses) ~ "Burana virus XXX",
                                grepl("Bronnoya virus", viruses) ~ "Bronnoya virus XXX",
                                grepl("Connecticut virus", viruses) ~ "Connecticut virus XXX",
                                grepl("Grand Arbaud virus", viruses) ~ "Grand Arbaud virus XXX",
                                grepl("Rondonia orthonairovirus", viruses) ~ "Rondonia orthonairovirus YYY",
                                grepl("Farallon virus", viruses) ~ "Farallon virus XXX",
                                grepl("Silverwater virus", viruses) ~ "Silverwater virus XXX",
                                grepl("Quaranjavirus sp.", viruses) ~ "Quaranjavirus sp. YYY",
                                grepl("Precarious point virus", viruses) ~ "Precarious point virus XXX",
                                grepl("Vinegar Hill virus", viruses) ~ "Vinegar Hill virus XXX",
                                grepl("Hunter Island virus", viruses) ~ "Hunter Island virus XXX",
                                grepl("Kuriyama virus", viruses) ~ "Kuriyama virus XXX",
                                grepl("Grotenhout virus", viruses) ~ "Grotenhout virus XXX",
                                grepl("Gomselga virus", viruses) ~ "Gomselga virus XXX",
                                grepl("Ngari virus", viruses) ~ "Ngari virus XXX",
                                grepl("Wanowrie virus", viruses) ~ "Wanowrie virus XXX",
                                grepl("Toyo virus", viruses) ~ "Toyo virus XXX",
                                grepl("Ngari virus", viruses) ~ "Ngari virus XXX",
                                grepl("Estero Real virus", viruses) ~ "Estero Real virus XXX",
                                grepl("Iftin tick virus", viruses) ~ "Iftin tick virus XXX",
                                grepl("Geran virus", viruses) ~ "Geran virus XXX",
                                grepl("Zambezi tick virus", viruses) ~ "Zambezi tick virus XXX",
                                grepl("Umea virus", viruses) ~ "Umea virus XXX",
                                grepl("Bonden virus", viruses) ~ "Bonden virus XXX",
                                grepl("Ubmeje virus", viruses) ~ "Ubmeje virus XXX",
                                grepl("Manawa virus", viruses) ~ "Manawa virus XXX",
                                grepl("Uukuniemi virus", viruses) ~ "Uukuniemi virus XXX",
                                grepl("Palma virus", viruses) ~ "Palma virus XXX",
                                grepl("Cencurut virus", viruses) ~ "Cencurut virus XXX",
                                grepl("Sunday Canyon virus", viruses) ~ "Sunday Canyon virus XXX",
                                grepl("Saphire II virus", viruses) ~ "Saphire II virus XXX",
                                grepl("New Kent County virus", viruses) ~ "New Kent County virus XXX",
                                grepl("Laurel Lake virus", viruses) ~ "Laurel Lake virus XXX",
                                grepl("Uukuniemi virus", viruses) ~ "Uukuniemi virus XXX",
                                grepl("Stavropol virus", viruses) ~ "Stavropol virus XXX",
                                grepl("Bronnoya virus", viruses) ~ "Bronnoya virus XXX",
                                grepl("Dandong tick virus", viruses) ~ "Dandong tick virus XXX",
                                grepl("Guyuan tick virus", viruses) ~ "Guyuan tick virus XXX",
                                grepl("Heihe tick virus", viruses) ~ "Heihe tick virus XXX",
                                grepl("Nanchang tick virus", viruses) ~ "Nanchang tick virus XXX",
                                grepl("Tonghua tick virus", viruses) ~ "Tonghua tick virus XXX",
                                grepl("Zhangye tick virus", viruses) ~ "Zhangye tick virus XXX",
                                grepl("Balambala tick virus", viruses) ~ "Balambala tick virus XXX",
                                grepl("Gakugsa tick virus", viruses) ~ "Gakugsa tick virus XXX",
                                grepl("Lone Star virus", viruses) ~ "Lone Star virus XXX",
                                grepl("Kuriyama virus", viruses) ~ "Kuriyama virus XXX",
                                grepl("Kumlinge virus", viruses) ~ "Kumlinge virus XXX",
                                grepl("Middelburg virus", viruses) ~ "Middelburg virus XXX",
                                grepl("Meaban virus", viruses) ~ "Meaban virus XXX",
                                grepl("Gadgets Gully virus", viruses) ~ "Gadgets Gully virus XXX",
                                grepl("Royal Farm virus", viruses) ~ "Royal Farm virus XXX",
                                grepl("Ngoye virus", viruses) ~ "Ngoye virus XXX",
                                grepl("Jiutai virus", viruses) ~ "Jiutai virus XXX",
                                grepl("Tyuleniy virus", viruses) ~ "Tyuleniy virus XXX",
                                grepl("Kama virus", viruses) ~ "Kama virus XXX",
                                grepl("Tyuleniy virus", viruses) ~ "Tyuleniy virus XXX",
                                grepl("Tacheng tick virus", viruses) ~ "Tacheng tick virus XXX",
                                grepl("Sikhote-Alin virus", viruses) ~ "Sikhote-Alin virus XXX",
                                grepl("Trinbago virus", viruses) ~ "Trinbago virus XXX",
                                grepl("Tacheng tick virus", viruses) ~ "Tacheng tick virus XXX",
                                grepl("Flaviviridae sp.", viruses) ~ "Flaviviridae sp. YYY",
                                grepl("Picornaviridae sp.", viruses) ~ "Picornaviridae sp. YYY",
                                grepl("Turkish sheep encephalitis virus", viruses) ~ "Turkish sheep encephalitis virus XXX",
                                grepl("Greek goat encephalitis virus", viruses) ~ "Greek goat encephalitis virus XXX",
                                grepl("Hubei tick virus", viruses) ~ "Hubei tick virus XXX",
                                grepl("Fennes virus", viruses) ~ "Fennes virus XXX",
                                grepl("Bulatov virus", viruses) ~ "Bulatov virus XXX",
                                grepl("Messner virus", viruses) ~ "Messner virus XXX",
                                grepl("Gerbovich virus", viruses) ~ "Gerbovich virus XXX",
                                grepl("Piguzov virus", viruses) ~ "Piguzov virus XXX",
                                grepl("Vovk virus", viruses) ~ "Vovk virus XXX",
                                grepl("Mackintosh virus", viruses) ~ "Mackintosh virus XXX",
                                grepl("Ed virus", viruses) ~ "Ed virus XXX",
                                grepl("Snesslinge virus", viruses) ~ "Snesslinge virus XXX",
                                grepl("Agalma virus", viruses) ~ "Agalma virus XXX",
                                grepl("Bolka virus", viruses) ~ "Bolka virus XXX",
                                grepl("Qualyub virus", viruses) ~ "Qualyub virus XXX",
                                grepl("Graso virus", viruses) ~ "Graso virus XXX",
                                grepl("Hepelivirales sp.", viruses) ~ "Hepelivirales sp. YYY",
                                grepl("Henan tick virus", viruses) ~ "Henan tick virus XXX",
                                grepl("Valo virus", viruses) ~ "Valo virus XXX",
                                grepl("Ganjam virus", viruses) ~ "Ganjam virus XXX",
                                grepl("Timbillica virus", viruses) ~ "Timbillica virus XXX",
                                grepl("Genoa virus", viruses) ~ "Genoa virus XXX",
                                grepl("Liman tick virus", viruses) ~ "Liman tick virus XXX",
                                 grepl("Canne point virus", viruses) ~ "Canne point virus XXX",
                                 grepl("Collins beach virus", viruses) ~ "Collins beach virus XXX",
                                 grepl("Fairfax lookout virus", viruses) ~ "Fairfax lookout virus XXX",
                                 grepl("Fairlight virus", viruses) ~ "Fairlight virus XXX",
                                grepl("Jump rock virus", viruses) ~ "Jump rock virus XXX",
                                 grepl("Manly virus", viruses) ~ "Manly virus XXX",
                                 grepl("Nadgee virus", viruses) ~ "Nadgee virus XXX",
                                 grepl("Old quarry swamp virus", viruses) ~ "Old quarry swamp virus XXX",
                                 grepl("Quarantine head virus", viruses) ~ "Quarantine head virus XXX",
                                grepl("Store beach virus", viruses) ~ "Store beach virus XXX",
                                 grepl("Store beach virus", viruses) ~ "Store beach virus XXX",
                                 grepl("Yambulla virus", viruses) ~ "Yambulla virus XXX",
                                 grepl("Ronne virus", viruses) ~ "Ronne virus XXX",
                                grepl("Fennes virus", viruses) ~ "Fennes virus XXX",
                                 grepl("Gerbovich virus", viruses) ~ "Gerbovich virus XXX",
                                 grepl("Bulatov virus", viruses) ~ "Bulatov virus XXX",
                                 grepl("Vovk virus", viruses) ~ "Vovk virus XXX",
                                grepl("Torhout tick virus", viruses) ~ "Torhout tick virus XXX",
                                grepl("Great Saltee virus", viruses) ~ "Great Saltee virus XXX",
                                grepl("Shanxi tick virus", viruses) ~ "Shanxi tick virus XXX",
                                grepl("Hubei tick virus", viruses) ~ "Hubei tick virus XXX",
                                grepl("Avalon Bres virus", viruses) ~ "Avalon Bres virus XXX",
                                grepl("Henan tick virus", viruses) ~ "Henan tick virus XXX",
                                grepl("Fort Crockett virus", viruses) ~ "Fort Crockett virus XXX"
                                
         )) 


# see if i can fix stragglers
unique(tick_virus_all_cleanvirus$fix_viruses)
tick_virus_all_cleanvirus[is.na(tick_virus_all_cleanvirus$fix_viruses),] # 553
View(tick_virus_all_cleanvirus[is.na(tick_virus_all_cleanvirus$fix_viruses),]) 

## okay let's see what happens when i get rid of generic species
#tick_virus_all_cleanvirus %>% 
#  filter(grepl("Reoviridae sp.", viruses) |

tick_virus_all_cleanvirus %>% 
  filter(grepl("associated", viruses)) # 129 associated


tick_virus_all %>% 
  filter(references == "Unpublished") %>% 
  tally() # 2933	

tick_virus_all %>% 
  filter(sample_type == "Culture") %>% 
  tally() # 1395	
```

##Step 8. Write csv of cleaned up 

There could be more fine tuning, but for now this is good enough for the purposes of the current paper frameworks
```{r}
#write.csv(tick_virus_all_cleanvirus, "tickviruses/clean/tick_viruses_clean_20240921.csv")
```

