---
title: "Analysis - BRT"
output: html_document
date: "2025-11-03"
---

**Manuscript title:** Ecological and socioeconomic factors associated with globally reported tick-borne viruses

**Corresponding author:** Samantha Sambado (sbsambado at ucsb.edu, ssambado at stanford.edu)

**Purpose**
To run boosted regression trees on cleaned data. Results from analysis will be in the final file: brt_result_presence_202501103_seed.rds

Other important flat files include: 
+ zover_matrix_formatted_preBRT_20251103.csv (cleaned data for figures/tables)
+ vinf_data_20251103.txt (presence model results for VIF figures)
+ brt_result_citation_20251103.rds (citation model results for figures)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

## upload necessary packages
library(readr)
library(tidyverse)
library(naniar)
library(gbm)
library(rsample)
library(ROCR)
library(InformationValue)
library(plotrix)

## upload necessary data
matrix_data <- read.csv("trait_matrix/clean/global_trait_matrix_20251031.csv") %>% dplyr::select(-X)
topvirus_data <- read.csv("tickviruses/clean/zover_toprecognized_slim_pubmed_20251103.csv" #"tickviruses/clean/zover_toprecognized_complete_pubmed _20240921.csv"
                          ) %>% dplyr::select(-X)

```

#Step 1. Merge top virus ZOVER data with global trait matrix
```{r}
# select only relevant information
data_slim <- topvirus_data %>% 
  group_by(admin) %>% 
  slice(1) %>% 
  dplyr::select(admin, iso_a3,tbv_presence, tbv_total, pubmedcitations_percounty_total) %>% 
  # merge trait matrix
  right_join(matrix_data, by = c("iso_a3")) %>% #, "admin"
  mutate_at(vars(tbv_presence, tbv_total, pubmedcitations_percounty_total), ~ replace_na(.,0)) %>% 
  rename(admin = admin.x) %>% 
  dplyr::select(-c(kopp_climatezone_ID,humpopcount_mean, shdi_mean, livestock_mean, vegdisim_mean, admin.y)) 

```

#Step 2. Format for BRT
```{r}
brt_data_formatted <- data_slim %>% 
  # make outcome a numeric
  mutate(tbv_presence = ifelse(tbv_presence>0,1,0)) %>% 
  # make characters factors
  mutate(kopp_climatezone_name = as.factor(kopp_climatezone_name),
         ivsa_chapter = as.factor(ivsa_chapter)) %>% 
  mutate_at(vars(#glw_livestockdensity_mean,
                 wb_livestockprod_index,
                 sedac_popcounts_sum,
                 wb_landarea_sqkm,
                 acled_conflictevents_total,
                 who_reportntd_mean,
                 #fao_pestusecapita_kmp,
                 wb_pestimportstot_mean), log) %>% 
  rename_with(~paste0("log_", .x),
              c(#glw_livestockdensity_mean,
                wb_livestockprod_index,
                sedac_popcounts_sum,
                wb_landarea_sqkm,
                acled_conflictevents_total,
                who_reportntd_mean,
                #fao_pestusecapita_kmp,
                wb_pestimportstot_mean)) %>% 
  # add +1 to citation count for log transformation
  mutate(log_pubmedcitations_percounty_total = log(pubmedcitations_percounty_total+1),
         log_fao_pestusecapita_kmp = log(fao_pestusecapita_kmp+1),
         log_glw_livestockdensity_mean = log(glw_livestockdensity_mean+1)) %>% 
  dplyr::select(-c(pubmedcitations_percounty_total,fao_pestusecapita_kmp,glw_livestockdensity_mean,# get rid of old var
                log_wb_pestimportstot_mean)) # <50 coverage


# check vars with 0 so log is transforming into -Inf
#unique(data_slim$fao_pestusecapita_kmp)
#unique(data_slim$glw_livestockdensity_mean)
#unique(data_slim$wb_livestockprod_index)
```

#Step 3. Data check for coverage & variance
```{r}
## check coverage visually
brt_data_formatted %>% 
  #dplyr::select(10:30) %>% 
  vis_miss()


## check 1. drop any variables with > 50% NAs
# calc the porportion of missing values in each column and keep only columns with < 50% missing values

varnums <- c(3:30)
var_check_na <- 2 + unname(which(c(colSums(is.na(brt_data_formatted[,varnums]))/nrow(brt_data_formatted)) < 0.5))
# columns 3-30 have more than 50% data present

## check 2. drop variables that lack variation
# calculate the proportion of the mode common (modal) value to assess how much variation is in a variable
mode.prop <- function(x) {
  ux <- unique(x[is.na(x)==FALSE]) # unique no-NA values
  tab <- tabulate(match(na.omit(x), ux)) # count occurrence of each unique variable
  max(tab)/length(x[is.na(x)==FALSE]) # prop of the mode (most frequent value)
}
var_check_prop <- 2 + unname(which(c(apply(brt_data_formatted[,varnums], 2, mode.prop)) < 0.95))
# columns 3-30 have more than 95% variation


## save as BRT dataset 
write.csv(brt_data_formatted, file = "zover_matrix_formatted_preBRT_20251103.csv", row.names = FALSE)

```

#Step 4. Hyperparameter tuning
Find best combination of tuning parameters
```{r}
zover_matrix_formatted <- read.csv("zover_matrix_formatted_preBRT_20251103.csv")


## step 1. format data for tbv presence model
data_tbvpresence <- zover_matrix_formatted %>% 
  dplyr::select(-c(admin, country, iso_a3, tbv_total, log_pubmedcitations_percounty_total)) %>% 
  # make characters factors
  mutate(kopp_climatezone_name = as.factor(kopp_climatezone_name),
         ivsa_chapter = as.factor(ivsa_chapter)) %>% 
  filter(!is.na(tbv_presence))

## step 2. define  grid of hyperparamters to tune BRT
hgrid <- expand.grid(n.trees = 5000, # base number of trees
                     interaction.depth = c(2, 3, 4), # tree complexity
                     shrinkage = c(0.01, 0.001, 0.0005), # learning rrate
                     n.minobsinnode = 4, # min observation in terminal node
                     seed = 1:10) # random seeds for multiple replicates

## make some adjustments to hgrid
# adjust number of trees for smaller shrinkage/learning rate (slower learning = more trees)
hgrid$n.trees <- ifelse(hgrid$shrinkage < 0.001, hgrid$n.trees*3, hgrid$n.trees)


# generate unique ID for each hyperparameter combo (excluding seed)
hgrid$id <- with(hgrid, paste(n.trees, interaction.depth, shrinkage, n.minobsinnode, sep = "_"))

# sort grid by ID and seed
hgrid <- hgrid[order(hgrid$id, hgrid$seed), ]

# add row index and numeric factor
hgrid$row <- 1:nrow(hgrid)
hgrid$id2 <- factor(as.numeric(hgrid$id))


## step 3. write function to assess each hyperparameter combo
hfit <- function(row, response) {
  
  # make new data
  model_df = data_tbvpresence
  
  # use rsample to split
  set.seed(hgrid$seed[row])
  split = initial_split(model_df, prop = 0.7, strata = "tbv_presence")
  
  # test and train data
  dataTrain = training(split)
  dataTest = testing(split)
  
  # ytest and ytrain
  yTrain = dataTrain$tbv_presence
  yTest = dataTest$tbv_presence

   
  ## fit BRT
  gbmOut = gbm(tbv_presence ~ . ,
               data = dataTrain,
               distribution = "bernoulli",
               n.trees = hgrid$n.trees[row],
               shrinkage = hgrid$shrinkage[row],
               interaction.depth = hgrid$interaction.depth[row],
               n.minobsinnode = hgrid$n.minobsinnode[row],
               cv.folds = 5,
               class.stratify.cv = TRUE,
               bag.fraction = 0.5,
               train.fraction = 1,
               n.cores = 1,
               verbose = F)
  
  # check which parameters were the best
  best.iter = gbm.perf(gbmOut, method = "cv", plot.it = FALSE)
  
  # predicted values
  preds = predict(gbmOut, dataTest, n.trees = best.iter, type = "response")
  
  # known values
  result = dataTest$tbv_presence
  
  ## BRT model metrics
  sen = InformationValue::sensitivity(result, preds)
  spec = InformationValue::specificity(result, preds)
  
  auc_train = gbm.roc.area(yTrain, predict(gbmOut, dataTrain, n.trees = best.iter, type = "response"))
  auc_test = gbm.roc.area(yTest, predict(gbmOut, dataTest, n.trees = best.iter, type = "response"))
  
  # print & save values
  print(paste("hpar row ", row, " done; test AUC is ", auc_test, sep = ""))
  
  return(list(best = best.iter,
              trainAUC = auc_train,
              testAUC = auc_test,
              sen = sen,
              spec = spec,
              wrow = row))
}


## step 4. run the function
hpars <- lapply(1:nrow(hgrid), function(i) hfit(i, response = "tbv_presence"))

## step 5. format results
# pull results
hresults <- data.frame(sapply(hpars, function(x) x$trainAUC),
                      sapply(hpars, function(x) x$testAUC),
                      sapply(hpars, function(x) x$sen),
                      sapply(hpars, function(x) x$spec),
                      sapply(hpars, function(x) x$wrow),
                      sapply(hpars, function(x) x$best))
# make column names
names(hresults) <- c("trainAUC", "testAUC", "sen", "spec", "row", "best")

## step 6. combine and save
hsearch <- merge(hresults, hgrid)

# write.csv 
#write.csv(hsearch, "hsearch_paramtuning_datasummary_20251103.csv", row.names = FALSE)
  
## step 7. identify which parameters are the best
# factor to make plotting easier
hsearch$shrinkage <- factor(hsearch$shrinkage)
lvl <-  rev(sort(unique(hsearch$shrinkage)))
hsearch$shrinkage <- factor(hsearch$shrinkage, levels = lvl); rm(lvl)
hsearch$interaction.depth <- factor(hsearch$interaction.depth)

# conver from wide to long
hsearch_long <- gather(hsearch, measure, value, testAUC:spec)

# chnage factors
hsearch_long$measure <- factor(hsearch_long$measure,
                               levels = c("testAUC", "sen", "spec"),
                               labels = c("test AUC", "sensitivity", "specificity"))

# okay now identify best parameters
hsearch_long %>% filter(measure == "test AUC") %>% arrange(desc(value))
# 18	0.9888112	3778	15000	3	5e-04	4	8
# n.tress = 15000
# interaction.dept = 3
# shrinkage/learning rate = 5e-04
# n.minobsinnode = 4
```

#Step 5. Presence model evaluation & prediction 
with fixed parameters across multiple random splits/seeds
```{r}
# n.tress = 15000
# interaction.dept = 3
# shrinkage/learning rate = 5e-04

## step 1. write function
brt_part_presence <- function(seed, response) {
  
  # make new data
  model_df = data_tbvpresence
  
  # use rsample to split
  set.seed(seed)
  split <- initial_split(model_df, prop = 0.7, strata = "tbv_presence")
  
  # test and train data
  dataTrain = training(split)
  dataTest = testing(split)
  
  # ytest and ytrain
  yTrain = dataTrain$tbv_presence
  yTest = dataTest$tbv_presence
  
  ## BRT model
  set.seed(1)
  
  gbmOut = gbm(tbv_presence ~ . ,
               data = dataTrain,
               distribution = "bernoulli",
               n.trees = 15000,
               shrinkage = 0.0005,
               interaction.depth = 3,
               n.minobsinnode = 4,
               cv.folds = 10,
               class.stratify.cv = TRUE,
               bag.fraction = 0.5,
               train.fraction = 1,
               n.cores = 1,
               verbose = 1)
  
  # check model performance with test data
  best.iter = gbm.perf(gbmOut, method = "cv", plot.it = FALSE)
  
  # predicted values
  preds <- predict(gbmOut, dataTest, n.trees = best.iter, type = "response")
  
  # known values
  result <- dataTest$tbv_presence
  
  ## model metrics
  sen = InformationValue::sensitivity(result, preds)
  spec = InformationValue::specificity(result, preds)
  
  auc_train = gbm.roc.area(yTrain, predict(gbmOut, dataTrain, n.trees = best.iter, type = "response"))
  auc_test = gbm.roc.area(yTest, predict(gbmOut, dataTest, n.trees = best.iter, type = "response"))
  
  ## variable relative importance
  bars <- summary(gbmOut, n.trees = best.iter, plotit = F)
  bars$rel.inf <- round(bars$rel.inf, 2)
  
  ## print & save values
  print(paste("Seed ", seed, " done; test AUC is ", auc_test, sep = ""))
  
  return(list(mod = gbmOut,
              best = best.iter,
              preds = preds, 
              trainAUC = auc_train,
              testAUC = auc_test,
              sen = sen,
              spec = spec,
              rinf = bars,
              traindata = dataTrain,
              testdata = dataTest,
              seed = seed))
}

## step 2. create stratified partitions to create an ensemble
smax = 101 # 100 model runs
brt_result_presence <- lapply(1:smax, function(x) brt_part_presence(seed = x, response = "tbv_presence"))


## step 3. save
saveRDS(brt_result_presence, "brt_result_presence_20251103.rds")

```

#Step 6. BRT metrics for presence model

###Test metrics results
```{r}
# get rid of runs that failed to converge
keep <- which(!is.na(sapply(brt_result_presence, function(x) x$testAUC)))
brt_result_presence <- brt_result_presence[keep]

### model metrics
## get mean AUC
mean(sapply(brt_result_presence, function(x) x$testAUC)) ## 0.8294715
std.error(sapply(brt_result_presence, function(x) x$testAUC)) # 0.004164542

## get mean sensitivity 
mean(sapply(brt_result_presence, function(x) x$sen)) # 0.4383938
std.error(sapply(brt_result_presence, function(x) x$sen)) #  0.01094209

## get mean specificity
mean(sapply(brt_result_presence, function(x) x$spec)) # 0.9044967
std.error(sapply(brt_result_presence, function(x) x$spec)) # 0.004977494


```


###VIF results
```{r}
# pull out variable relative influence
vinf <- lapply(brt_result_presence, function(x) x$rinf)
response_vinf <- do.call(rbind, vinf)

## aggreagte mean, se and rel.inf
vinf_data <- data.frame(aggregate(rel.inf ~ var, data = response_vinf , mean),
                        aggregate(rel.inf ~ var, data = response_vinf, std.error)["rel.inf"],
                        aggregate(rel.inf ~ var, data = response_vinf, var)["rel.inf"])
names(vinf_data) <- c("var", "rel.inf", "rse", "rvar")

vinf_data <- vinf_data[order(vinf_data$rel.inf, decreasing = T),]

## calculate vif metrics
mean(vinf_data$rel.inf) # 4.166828
var(vinf_data$rel.inf) # 29.23048

# rnak 
vinf_data$response_rank <- 1:nrow(vinf_data)


# save as table to make figures
#write.table(vinf_data, file = "vinf_data_20251103.txt", sep = ",", quote = FALSE, row.names = F)
```

#Step 7. Citation effort model
Now run poisson BRT with citation as outcome to see if reported TBV is a reflection of sampling effort
```{r}
##step 1. format citation data
zover_matrix_formatted_presence <- data_slim %>% 
  # make characters factors
  mutate(kopp_climatezone_name = as.factor(kopp_climatezone_name),
         ivsa_chapter = as.factor(ivsa_chapter)) %>% 
  mutate_at(vars(#glw_livestockdensity_mean,
                 wb_livestockprod_index,
                 sedac_popcounts_sum,
                 wb_landarea_sqkm,
                 acled_conflictevents_total,
                 who_reportntd_mean,
                 #fao_pestusecapita_kmp,
                 wb_pestimportstot_mean), log) %>% 
  rename_with(~paste0("log_", .x),
              c(#glw_livestockdensity_mean,
                wb_livestockprod_index,
                sedac_popcounts_sum,
                wb_landarea_sqkm,
                acled_conflictevents_total,
                who_reportntd_mean,
                #fao_pestusecapita_kmp,
                wb_pestimportstot_mean)) %>% 
  # add +1 to citation count for log transformation
  mutate(log_pubmedcitations_percounty_total = log(pubmedcitations_percounty_total+1),
         log_fao_pestusecapita_kmp = log(fao_pestusecapita_kmp+1),
         log_glw_livestockdensity_mean = log(glw_livestockdensity_mean+1)) %>% 
  mutate(log_pubmedcitations_percounty_total = as.integer(log_pubmedcitations_percounty_total)) %>% 
  dplyr::select(-c(pubmedcitations_percounty_total,fao_pestusecapita_kmp,glw_livestockdensity_mean,# get rid of old var
                log_wb_pestimportstot_mean)) # <50 coverage




## step 2. format data for citation model
data_citation <- zover_matrix_formatted_presence%>% 
  dplyr::select(-c(admin, country, iso_a3, tbv_total, tbv_presence)) %>% 
  # make characters factors
  mutate(kopp_climatezone_name = as.factor(kopp_climatezone_name),
         ivsa_chapter = as.factor(ivsa_chapter)) %>% 
  filter(!is.na(log_pubmedcitations_percounty_total))


## step 3. write function
brt_part_citation <- function(seed, response) {
  
  # make new data
  model_df = data_citation
  
  # use rsample to split
  set.seed(seed)
  split <- initial_split(model_df, prop = 0.7, strata = "log_pubmedcitations_percounty_total")
  
  # test and train data
  dataTrain = training(split)
  dataTest = testing(split)
  
  
  ## BRT model
  set.seed(1)
  
  gbmOut = gbm(log_pubmedcitations_percounty_total ~ . ,
               data = dataTrain,
               distribution = "poisson",
               n.trees = 15000,
               shrinkage = 0.0005,
               interaction.depth = 3,
               n.minobsinnode = 4,
               cv.folds = 10,
               #class.stratify.cv = TRUE,
               bag.fraction = 0.5,
               train.fraction = 1,
               n.cores = 1,
               verbose = 1)
  
  # check model performance with test data
  best.iter = gbm.perf(gbmOut, method = "cv", plot.it = FALSE)
  
  # predicted values
  preds <- predict(gbmOut, dataTest, n.trees = best.iter, type = "response")
  
  # known values
  result <- dataTest$log_pubmedcitations_percounty_total
  
  ## model metrics
  rmse <- sqrt(mean((result - preds)^2))
  mae <- mean(abs(result - preds))
  correlation <- cor(result, preds, method = "spearman")
  
  
  # make structured prediciton dataframe
  pred_df <- data.frame(
    observed = result,
    predicted = preds,
    seed = seed)
  
  ## print & save values
  print(paste("Seed ", seed, " done; RMSE is ", rmse, sep = ""))
  
  return(list(best = best.iter,
              rmse = rmse,
              mae = mae,
              correlation = correlation,
              seed = seed,
              preds = preds, 
              testdata = dataTest,
              predict = pred_df))
}

## step 4. create stratified partitions to create an ensemble
smax = 101
brt_result_citation <- lapply(1:smax, function(x) brt_part_citation(seed = x, response = "log_pubmedcitations_percounty_total"))


## step 5. save
#saveRDS(brt_result_citation, "brt_result_citation_20251103.rds")



### step 6. BRT model metrics summary
# drop failed runs
keep <- which(!is.na(sapply(brt_result_citation, function(x) x$rmse)))
brt_result_citation <- brt_result_citation[keep]

## summary metrics
rmse_mean <- mean(sapply(brt_result_citation, function(x) x$rmse)) # 4.309495
rmse_se <- std.error(sapply(brt_result_citation, function(x) x$rmse)) # 0.02131648

mae_mean <- mean(sapply(brt_result_citation, function(x) x$mae)) # 3.113222
mae_se <- std.error(sapply(brt_result_citation, function(x) x$mae)) # 0.01619675

cor_mean <- mean(sapply(brt_result_citation, function(x) x$correlation)) # 0.5043058
cor_se <- std.error(sapply(brt_result_citation, function(x) x$correlation)) # 0.007225013
                                           

## poisson moderate predictive power for citation effort using the same predictors from bernoulli model. the correlation is 0.49 and error is relatively large (3-4 citations off on avera)... predictors of tbv presence are note merely a reflection of citation effort, if they were the poisson model would likely have stronger predictor performance.. eco predictos may genuinely related to where viruses exist, not just where people study them.. my classification model is not simply pricking up citaion biases bc the same predictors are less effective at predicting citation effort
```

