---
title: "3_ZOVEROutcomes_GlobalTBVs"
output: html_document
date: "2024-09-21"
---

**Purpose** 
To further refine ZOVER data for the purpose of creating a dataset that represents only the top ~15 tick-borne viruses to be merged with global trait matrix

**Steps**

1) Find what are the top reported TBVs on ZOVER (n > 30)
2) Fix the names of the top reported TBVs
3) Filter to top ~15 viruses between 1990-2023
4) Aggregated counts per country & assign 1/0 per country
5) Slim the dataset to just relevant ZOVER metadata 
6) Merge aggregated country counts with relevant ZOVER metadata
7) Run some basic summaries for Methods
8) Add citation counts using easyPubMed
9) Aggregate citation count per country
10) Give each country a citation weight
11) Write csv for top reported tick-borne viruses

###Final CSV: zover_toprecognized_complete_20240921.csv
^^ this will be data that is merged with global trait matrix

Set up

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

## upload necessary packages
library(ggplot2)
library(tidyverse)
library(readr)
library(janitor)
library(rnaturalearth)
library(easyPubMed)

## upload necessary data
# clean-ish ZOVER data for all viruses
zover_all <- read.csv("tickviruses/clean/tick_viruses_clean_20240921.csv")

# 
worldmap <- ne_countries(scale = "large", type = "countries", returnclass = "sf") %>% 
```

##Step 1. Find what are the top reported TBVs on ZOVER (n > 30)

```{r}

## first identify the current 30
zover_all %>% 
  group_by(viral_family,fix_viruses) %>% 
  tally() %>% 
  arrange(desc(n)) %>% 
  # omit names that are meaningless
  filter(fix_viruses != "NA" &
           fix_viruses != "Rhabdoviridae sp. YYY" &
           viral_family != "N/A" &
           fix_viruses != "Flaviviridae sp. YYY" &
           fix_viruses != "Nairoviridae sp. YYY" &
           fix_viruses != "Peribunyaviridae sp. YYY" &
           fix_viruses != "Orthomyxoviridae sp. YYY" &
           fix_viruses != "Phenuiviridae sp. YYY" &
           fix_viruses != "Bocaparvovirus" &
           fix_viruses != "Phlebovirus XXX" &
           n >30) 

```

##Step 2. Fix the names of the top reported TBVs
```{r}
## fix names of top ~20 viruses
zover_all <- zover_all %>% 
  mutate(fix_viruses = case_when(fix_viruses == "Deer tick virus XXX" ~ "Deer tick virus",
                                 fix_viruses == "Crimean-Congo hemorrhagic fever"~ "Crimean-Congo Hemorrhagic fever",
                                 fix_viruses == "Severe fever with thrombocytopenia virus"~ "Severe fever with thrombocytopenia syndrome",
                                 fix_viruses == "Jingmen tick virus XXX"~ "Jingmen tick virus",
                                 fix_viruses == "African swine fever"~ "African swine fever virus",
                                 fix_viruses == "Dabieshan Tick"~ "Dabieshan tick virus",
                                 fix_viruses == "Wuhan tick virus"~ "Wuhan tick virus",
                                 fix_viruses == "Alongshan virus XXX"~ "Alongshan virus",
                                 fix_viruses == "Lihan tick virus XXX"~ "Lihan tick virus",
                                 fix_viruses == "Yanggou tick virus XXX"~ "Yanggou tick virus",
                                 fix_viruses == "Mukawa virus XXX"~ "Mukawa virus",
                                 fix_viruses == "Bole tick XXX"~ "Bole tick virus",
                                 fix_viruses == "Haseki tick virus XXX"~ "Haseki tick virus",
                                 fix_viruses == "Taishun Tick virus XXX"~ "Taishun tick virus",
                                 fix_viruses == "Kyasanur Forest disease virus  XXX"~ "Kyasanur Forest disease virus",
                                 fix_viruses == "Guangxi tick virus XXX"~ "Guangxi tick virus",
                                 fix_viruses == "Dugbe orthonairovirus XXX"~ "Dugbe orthonairovirus",
                                 fix_viruses == "Huaiyangshan virus XXX"~ "Huaiyangshan virus",
                                 fix_viruses == "Huangpi tick Virus"~ "Huangpi tick virus", 
                                 fix_viruses == "Alkhumra hemorrhagic fever virus XXX" ~ "Alkhumra hemorrhagic fever virus",
                                 fix_viruses == "Bhanja virus XXX" ~ "Bhanja virus",
                                 fix_viruses == "Louping ill virus XXX" ~ "Louping ill virus",
                                 fix_viruses == "Omsk hemorrhagic fever virus XXX" ~ "Omsk hemorrhagic fever virus",
                                 TRUE ~ fix_viruses))  

#zover_all %>% filter(grepl("Saw", fix_viruses)) # helps find what is reported in column

```

##Step 3. Filter to top ~15 viruses between 1990-2023

(based on Mansfield 2017, Lippi 2022, de la Fuente 2023)

+ African swine fever
+ Alkhumra hemorrhagic fever virus
+ Bhanja virus
+ Bourbon virus
+ Colorado tick fever
+ Crimean-Congo hemorrhagic fever
+ Deer tick virus
+ Heartland virus
+ Jingmen tick virus
+ Kyasanur Forest disease virus
+ Louping ill virus
+ Lumpy skin disease
+ Nairobi sheep disease virus
+ Omsk hemorrhagic fever virus
+ Powassan virus
+ Sawgrass virus
+ Severe fever with thrombocytopenia virus
+ Tick-borne encephalitis virus

```{r}
zover_toprecognized <- zover_all %>% # 7419 to 4213
  filter(grepl("Heart",fix_viruses) |
          grepl("Bourbon",fix_viruses) |
          grepl("Severe",fix_viruses) |
          grepl("Crimean",fix_viruses) |
          grepl("Powassan",fix_viruses) |
          grepl("Deer tick",fix_viruses) |
          grepl("Kyasanur",fix_viruses) |
          grepl("Alkhumra",fix_viruses) |
          grepl("African swine",fix_viruses) |
          grepl("Colorado",fix_viruses) |
          grepl("Louping",fix_viruses) |
          grepl("Nairobi",fix_viruses) |
          grepl("Omsk",fix_viruses) |
          grepl("Saw",fix_viruses) |
          grepl("Bhanja",fix_viruses) |
          grepl("Lumpy",fix_viruses) |
          grepl("Tick-borne enceph", fix_viruses) |
          grepl("Jingmen",fix_viruses)) %>% 
  ### THINK ABOUT YEAR CUTOFF -- CHANGE HERE IF MAKE DIFFERENT DECISION ###
  filter(collection_year > 1989) 
# got rid of Ganjam

zover_toprecognized %>% 
  group_by(sample_type) %>% 
  tally()


```


##Step 4. Aggregated counts per country & assign 1/0 per country

```{r}
unique(zover_all$sampling_country)

zover_toprecognized_countryaggregate <- zover_toprecognized %>% 
  group_by(sampling_country) %>% 
  tally() %>% 
  # make sure country names match rnaturalearth
  rename(admin = sampling_country) %>% 
  right_join(worldmap, by = "admin") %>% 
  mutate(iso_a3 = case_when(iso_a3 == "-99" ~ admin, # these are a handful of places like Naval Base Guantanamo but want to make sure they get merged on to final product so will keep their admin name
                            TRUE ~ iso_a3)) %>% 
  # assign binomial count
  mutate(tbv_presence = case_when(n > 0 ~ 1,
                                  is.na(n) ~ 0),
         n = case_when(is.na(n) ~ 0,
                       TRUE ~ n)) %>% 
  rename(tbv_total = n) %>% 
  dplyr::select(admin, iso_a3, tbv_presence, tbv_total) 
```

##Step 5. Slim the dataset to just relevant ZOVER metadata 

```{r}
zover_toprecognized_slim <- zover_toprecognized %>% 
  dplyr::select(sampling_country, fix_viruses, # most important info
                viral_family, type, # good info
                continent,year_agg,  # for visualizations
                from_tick, tick_genus, # for networks
                references, sample_type,viruses, iso3c) %>% # info may be helpful for filtering
  rename(admin = sampling_country)
```

##Step 6. Merge aggregated country counts with relevant ZOVER metadata

```{r}
zover_toprecognized_complete <- zover_toprecognized_countryaggregate %>% 
  right_join(zover_toprecognized_slim, by = "admin") # keep all rows in zover_toprecog_zlim
      
```

##Step 7. Run some basic summaries for Methods
```{r}
zover_toprecognized_complete %>% 
  group_by(viral_family) %>% 
  tally() # 8 virual families

zover_toprecognized_complete %>% 
  group_by(fix_viruses,type) %>% 
  tally()  %>% 
  filter(str_detect(type, "DNA"))


zover_toprecognized_complete %>% 
  group_by(fix_viruses,type) %>% 
  tally()  %>% 
  filter(str_detect(type, "ssRNA"))


zover_toprecognized_complete %>% 
  group_by(tick_genus) %>% 
  tally() %>% 
  filter(tick_genus != "Argas" &
           tick_genus != "N/A")


zover_toprecognized_complete %>% 
  group_by(fix_viruses) %>% 
  tally() 


```

##Step 8. Add citation counts using easyPubMed

```{r}

## check which version gets more hits
get_pubmed_ids("Tick-borne encephalitis")$Count
# African swine 4125 / African swine fever 3474
# Alkhumra hemorrhagic 34 / Alkhumra hemorrhagic fever 32
# Bourbon 724 / Bourbon virus 43
# Colorado tick 581 / Colorado tick fever 288
# Crimean-congo hemorrhagic 2131 / Crimean-congo hemorrhagic fever 2119
# Deer tick 10316 / Deer tick virus 1590
# Heartland 3364 / Heartland virus 111
# Jingmen tick 56 / Jingmen tick virus 56
# Kyasanur forest 298 / Kyasanur forest disease 297
# Lumpy skin 711 / Lumpy skin disease 681
# Nairobi sheep 399 / Nairobi sheep disease 226
# Powassan 411 / Powassan virus 4211
# Severe fever with thrombocytopenia syndrome 1983 / Severe fever with thrombocytopenia syndrome virus 1367
# Tick-borne encephalitis 6561 / Tick-borne encephalitis virus 4954 


## pull list of names
toprecognized_list <- zover_toprecognized_complete %>% 
  group_by(fix_viruses) %>% 
  slice(1) %>% 
  dplyr::select(fix_viruses) %>% 
  # fix names to match NCBI Taxonomy Browser
  # https://www.ncbi.nlm.nih.gov/Taxonomy/Browser/wwwtax.cgi
  mutate(fix_viruses = case_when(fix_viruses == "African swine fever virus" ~ "African swine",
                                 fix_viruses == "Alkhumra hemorrhagic fever virus" ~ "Alkhumra hemorrhagic",
                                 fix_viruses == "Colorado tick fever" ~ "Colorado tick fever",
                                 fix_viruses == "Crimean-Congo Hemorrhagic fever" ~ "Crimean-Congo hemorrhagic fever",
                                 fix_viruses == "Deer tick virus" ~ "Deer tick virus",
                                 fix_viruses == "Kyasanur Forest disease virus" ~ "Kyasanur forest disease",
                                 fix_viruses == "Lumpy skin disease" ~ "Lumpy skin disease",
                                 fix_viruses == "Nairobi sheep disease virus" ~ "Nairobi sheep disease",
                                 fix_viruses == "Severe fever with thrombocytopenia syndrome" ~ "Severe fever with thrombocytopenia syndrome",
                                 fix_viruses == "Tick-borne encephalitis virus" ~ "Tick-borne encephalitis",
                                 TRUE ~ fix_viruses))


## collect citations per virus species
# one way of doing this
# cites = c()
# for(i in 1:length(toprecognized_list$fix_viruses)) {
#   counts = as.numeric(as.character(get_pubmed_ids(gsub("_","-", toprecognized_list$fix_viruses[i]))$Count))
#   cites[i] = counts
#   print(paste(i,"/", nrow(toprecognized_list)))
# }

# looks for variations in virus or viral
vcites = c()
for(i in 1:length(toprecognized_list$fix_viruses)) {
  x = gsub("_","-", toprecognized_list$fix_viruses[i])
  x = paste("(",x,")",sep="")
  x = paste(x,"AND (virus OR viral)")
  counts = as.numeric(as.character(get_pubmed_ids(x)$Count))
  vcites[i]=counts
  print(paste(i,"/", nrow(toprecognized_list)))
}


# turn into dataframe
# virus citations
vcites_df <- as.data.frame(vcites)
# bind names with virus citations
toprecognized_pubmedcitations <- cbind(toprecognized_list, vcites_df)

```

##Step 9. Aggregate citation count per country
```{r}

zover_toprecognized_countryaggregate_pubmedcitations <- 
zover_toprecognized_complete %>% 
  mutate(fix_viruses = case_when(fix_viruses == "African swine fever virus" ~ "African swine",
                                 fix_viruses == "Alkhumra hemorrhagic fever virus" ~ "Alkhumra hemorrhagic",
                                 fix_viruses == "Colorado tick fever" ~ "Colorado tick fever",
                                 fix_viruses == "Crimean-Congo Hemorrhagic fever" ~ "Crimean-Congo hemorrhagic fever",
                                 fix_viruses == "Deer tick virus" ~ "Deer tick virus",
                                 fix_viruses == "Kyasanur Forest disease virus" ~ "Kyasanur forest disease",
                                 fix_viruses == "Lumpy skin disease" ~ "Lumpy skin disease",
                                 fix_viruses == "Nairobi sheep disease virus" ~ "Nairobi sheep disease",
                                 fix_viruses == "Severe fever with thrombocytopenia syndrome" ~ "Severe fever with thrombocytopenia syndrome",
                                 fix_viruses == "Tick-borne encephalitis virus" ~ "Tick-borne encephalitis",
                                 TRUE ~ fix_viruses)) %>% 
left_join(toprecognized_pubmedcitations, by = "fix_viruses") %>% 
  group_by(admin) %>% 
  summarise(pubmedcitations_percounty_total = sum(vcites))
```

##Step 10. Give each country a citation weight

Merge pubmed citation list with complete zover top recognized list
```{r}
zover_toprecognized_complete_pubmed <- zover_toprecognized_complete %>% 
  left_join(zover_toprecognized_countryaggregate_pubmedcitations, by = "admin")
```

##Step 11. Write csv for top reported tick-borne viruses
```{r}
#write.csv(zover_toprecognized_complete_pubmed , file = "tickviruses/clean/zover_toprecognized_complete_pubmed _20240921.csv")
```




Create a dataframe just for BRT analysis

