---
title: "Results for Data Summaries and Tables"
output: html_document
date: "2025-11-03"
---

**Manuscript title:** Ecological and socioeconomic factors associated with globally reported tick-borne viruses

**Corresponding author:** Samantha Sambado (sbsambado at ucsb.edu, ssambado at stanford.edu)

**Purpose**
To re-create results from text and tables in manuscript.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

##upload necessary packages
library(ggplot2)
library(tidyverse)

##upload neccessary data
zover_toprecognized_slim_pubmed <- read.csv("tickviruses/clean/zover_toprecognized_slim_pubmed_20251103.csv")
```

#Main Text

##Results

###Dataset description
```{r}

## how many countries have reported tbv of this list?
zover_toprecognized_complete_pubmed %>% 
  group_by(admin) %>% 
  slice(1) %>% 
  filter(tbv_presence == 1) # 59 countries


## which country has the most type of tbvs?
zover_toprecognized_complete_pubmed %>% 
  group_by(admin) %>% 
  summarise(count = n_distinct(fix_viruses)) %>% 
  arrange(desc(count)) #%>% 
  #filter(count == 1)

## which country had total most tbv?
zover_toprecognized_complete_pubmed %>% 
  group_by(admin) %>% 
  slice(1) %>% 
  summarise(count = sum(tbv_total)) %>% 
  arrange(desc(count)) 

## which was the most commonly reported tbv?
zover_toprecognized %>% 
  group_by(fix_viruses) %>% 
  tally() %>% 
  arrange(desc(n))
```

###BRT results
(TBV presence model)
```{r}

brt_result_presence <- readRDS("brt_result_presence_20251103.rds")

###Metrics
# get rid of runs that failed to converge
keep <- which(!is.na(sapply(brt_result_presence, function(x) x$testAUC)))
brt_result_presence <- brt_result_presence[keep]

### model metrics
## get mean AUC
mean(sapply(brt_result_presence, function(x) x$testAUC)) ## 0.8294715
std.error(sapply(brt_result_presence, function(x) x$testAUC)) # 0.004164542

## get mean sensitivity 
mean(sapply(brt_result_presence, function(x) x$sen)) # 0.4383938
std.error(sapply(brt_result_presence, function(x) x$sen)) #  0.01094209

## get mean specificity
mean(sapply(brt_result_presence, function(x) x$spec)) # 0.9044967
std.error(sapply(brt_result_presence, function(x) x$spec)) # 0.004977494


###VIF results
# pull out variable relative influence
vinf <- lapply(brt_result_presence, function(x) x$rinf)
response_vinf <- do.call(rbind, vinf)

## aggreagte mean, se and rel.inf
vinf_data <- data.frame(aggregate(rel.inf ~ var, data = response_vinf , mean),
                        aggregate(rel.inf ~ var, data = response_vinf, std.error)["rel.inf"],
                        aggregate(rel.inf ~ var, data = response_vinf, var)["rel.inf"])
names(vinf_data) <- c("var", "rel.inf", "rse", "rvar")

vinf_data <- vinf_data[order(vinf_data$rel.inf, decreasing = T),]

## calculate vif metrics
mean(vinf_data$rel.inf) # 4.166828
var(vinf_data$rel.inf) # 35.6388

# rnak 
vinf_data$response_rank <- 1:nrow(vinf_data)

```

(Citation model)
```{r}
# upload model results
brt_result_citation <- readRDS("brt_result_citation_20251103.rds")

# drop runs that failed to converge
keep <- which(!is.na(sapply(brt_result_citation, function(x) x$rmse)))
brt_result_citation <- brt_result_citation[keep]

## summary metrics
rmse_mean <- mean(sapply(brt_result_citation, function(x) x$rmse)) # 4.309495
rmse_se <- std.error(sapply(brt_result_citation, function(x) x$rmse)) # 0.02131648

mae_mean <- mean(sapply(brt_result_citation, function(x) x$mae)) # 3.113222
mae_se <- std.error(sapply(brt_result_citation, function(x) x$mae)) # 0.01619675

cor_mean <- mean(sapply(brt_result_citation, function(x) x$correlation)) # 0.5043058
cor_se <- std.error(sapply(brt_result_citation, function(x) x$correlation)) # 0.007225013
```

#Supplementary Tables

###Table 1
tick-borne viruses per country
```{r}
zover_topviruses <- read.csv("tickviruses/clean/zover_toprecognized_complete_pubmed _20240921.csv") %>% dplyr::select(-X)

table_s_1 <- zover_topviruses %>% 
  group_by(admin, iso_a3, fix_viruses) %>% 
  tally() %>% 
  mutate(fix_viruses = case_when(fix_viruses == "Lumpy skin disease" ~ "LSD",
                                        fix_viruses == "African swine fever virus" ~ "ASF",
                                        fix_viruses == "Colorado tick fever" ~ "CTF",
                                        fix_viruses == "Crimean-Congo Hemorrhagic fever" ~ "CCHF",
                                        fix_viruses == "Severe fever with thrombocytopenia syndrome" ~ "SFTS",
                                        fix_viruses == "Nairobi sheep disease virus" ~ "NSD",
                                        fix_viruses == "Sawgrass virus" ~ "SGV",
                                        fix_viruses == "Heartland virus" ~ "HLV",
                                        fix_viruses == "Bhanja virus" ~ "BHV",
                                        fix_viruses == "Bourbon virus" ~ "BOV",
                                        fix_viruses == "Jingmen tick virus" ~ "JTV",
                                        fix_viruses == "Powassan virus" ~ "PWV",
                                        fix_viruses == "Tick-borne encephalitis virus" ~ "TBE",
                                        fix_viruses == "Alkhumra hemorrhagic fever virus" ~ "AHF",
                                        fix_viruses == "Kyasanur Forest disease virus" ~ "KFD",
                                        fix_viruses == "Deer tick virus" ~ "DTF",
                                        fix_viruses == "Omsk hemorrhagic fever virus" ~ "OHF",
                                        fix_viruses == "Louping ill virus" ~ "LIV",
                                        TRUE ~ fix_viruses)) %>% 
    pivot_wider(names_from = fix_viruses, values_from = n) %>% 
  mutate_all(~ifelse(is.na(.), 0, .)) %>% 
  filter(admin != "ND") %>% 
  ungroup()

#write.table(table_s_1, file = "table_s_1.txt", sep = ",")

```

###Table 2
variable name, coverage, range
```{r}
brt_matrix_data <- read.csv("zover_matrix_formatted_preBRT_20251103.csv")

table_s_2 <- brt_matrix_data %>%
  dplyr::select(where(is.numeric)) %>%            
  pivot_longer(cols = everything(), 
               names_to = "variable", 
               values_to = "value") %>%     
  group_by(variable) %>%
  summarise(
    coverage = round(mean(!is.na(value)) * 100,1),
    min = round(min(value, na.rm = TRUE),0),
    max = round(max(value, na.rm = TRUE),0),
    .groups = "drop") %>% 
  mutate(range = paste(min, max, sep = "-")) %>% 
  mutate(var_pretty = case_when(variable== "kopp_climatezone_name" ~ "Climate Zone", 
                              variable== "wb_gini_index" ~ "Gini Index", 
                              variable== "log_sedac_popcounts_sum" ~ "Pop Counts", 
                              variable== "log_fao_pestusecapita_kmp" ~ "Pesticide Use", 
                              variable== "log_wb_landarea_sqkm" ~ "Land Area", 
                              variable== "log_wb_livestockprod_index" ~ "Livestock Index", 
                              variable== "ivsa_chapter" ~ "IVSA", 
                              variable== "log_acled_conflictevents_total" ~ "Conflict (total)", 
                              variable== "wb_povertyline_pct" ~ "Below Poverty Line", 
                              variable== "acled_conflictpopexposure_pct" ~ "Conflict (% exposed)",
                              variable== "log_glw_livestockdensity_mean" ~ "Livestock Density",
                              variable== "wb_gradvetsforestag_pct" ~ "% Ag/Vet/Forestry Grads",
                              variable== "wb_urbanpop_pct" ~ "Urban Pop",
                              variable== "wb_adultlit_rate" ~ "Adult Literacy",
                              variable== "log_who_reportntd_mean" ~ "NTD Reports",
                              variable== "wb_employedag" ~ "Ag Employment",
                              variable== "wb_gradhealthwelf_pct" ~ "% Healthcare/Welfare Grads",
                              variable== "wb_expendhealth_pct" ~ "Health Spending",
                              variable== "grdi_shdi_mean" ~ "Social Vulnerability",
                              variable== "wwf_ecoreg_total" ~ "Total Ecoregions",
                              variable== "wb_forested_pct" ~ "Forested",
                              variable== "wb_expendedu_pct" ~ "Education Spending",
                              variable== "ee_vegdisim_mean" ~ "Habitat Fragmentation",
                              variable== "wb_accesselect_pct" ~ "Access to Electricity",
                              variable == "tbv_presence" ~ "Presence of TBV",
                              variable == "tbv_total" ~ "Total TBV",
                              variable == "log_pubmedcitations_percounty_total" ~ "Total Citations",
                              TRUE ~ NA)) %>%
  dplyr::select(-variable) %>% 
  dplyr::select(var_pretty, coverage, range)

#write.table(table_s_2, file = "table_s_2.txt", sep = ",")
 

table_s_2 %>% 
  arrange(desc(coverage))
```

###Table 4
summary of variable relative influence
```{r}
vinf_data <- read.delim("vinf_data_20251103.txt",sep = ",")

# mean and variance of all vars
table_s_3 <- vinf_data %>% 
  group_by(var) %>% 
  summarise(across(where(is.numeric), ~ round(., 2))) %>% 
  arrange(desc(rel.inf))


#write.table(table_s_3, file = "table_s_3.txt", sep = ",")


# mean and standard error of all vars
# pull out variable relative influence
vinf <- lapply(brt_result_presence, function(x) x$rinf)
response_vinf <- do.call(rbind, vinf)
mean(vinf_data$rel.inf)
var(vinf_data$rel.inf) # 35.6388 or 
# report standard error?
vinf_data %>% 
  summarise(mean_importance = mean(rel.inf), se_importance = sd(rel.inf) / sqrt(n()))
```