---
title: "Figures"
output: html_document
date: "2025-11-04"
---

**Manuscript title:** Ecological and socioeconomic factors associated with globally reported tick-borne viruses

**Corresponding author:** Samantha Sambado (sbsambado at ucsb.edu, ssambado at stanford.edu)

**Purpose**
To run boosted regression trees on cleaned data.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

##upload necessary packages
library(ggplot2)
library(tidyverse)
library(ggpubr)
library(rnaturalearth)
library(patchwork)
library(scales)
library(RColorBrewer)

##upload necessary data
worldmap <- ne_countries(scale = "large", type = "countries", returnclass = "sf")
worldmap_slim <- worldmap %>% dplyr::select(admin, iso_a3)
brt_matrix_data <- read.csv("zover_matrix_formatted_preBRT_20251103.csv")
top <- read.csv("tickviruses/clean/zover_toprecognized_complete_pubmed _20240921.csv")
vinf_data <- read.delim("vinf_data_20251103.txt", sep = ",")
brt_presence_results <- readRDS("brt_result_presence_20251103.rds")
hsearch_results <- read.csv("hsearch_paramtuning_datasummary_20251103.csv")
all_tbvs <- read.csv("tickviruses/clean/zover_allviruses_20240921.csv")
```

#Main Text

##Figure 2
Zover database summary

###part a
```{r}
## upload data
world_matrix_data <- left_join(worldmap_slim , brt_matrix_data%>% dplyr::select(-admin), by = "iso_a3")


## wrangle data
world_matrix_clean <- world_matrix_data %>% 
  mutate(tbv_presence = replace_na(as.numeric(tbv_presence), 0),
         tbv_presence = factor(tbv_presence, levels = c(0,1), labels = c("Absent", "Present")),
         tbv_total = replace_na(as.numeric(tbv_total), 0),
         tbv_total_bin = cut(tbv_total, 
                             breaks = c(0, 1, 10, 50, 100, 200, 1000, Inf),
                             labels = c("0", "1-9", "10-49", "50-99", "100-199", "200-999", "1000+"),
                             right = FALSE),
         tbv_total_bin = factor(tbv_total_bin, levels = rev(levels(tbv_total_bin))))

## make color bin
bin_colors <- c("0"  = "#ECE9FC", "1-9"  = "#CABFF4", "10-49"  = "#A495EB", "50-99"  = "#8169E4",
                "100-199"= "#7B50C8","200-999"= "#783CA9", "1000+"  = "#77278B")  

## plot it

fig_2a <- ggplot() +
  geom_sf(data = world_matrix_clean, aes(fill = tbv_total_bin), color = NA) +
  geom_sf(data = world_matrix_clean, fill = NA, color = "black", size = 0.2) +
  scale_fill_manual(values = bin_colors, name = "Total Records", drop = FALSE,
    guide = guide_legend(direction = "vertical",
                         keyheight = unit(0.6, "cm"),
                         keywidth = unit(0.5, "cm"),
                         title.position = "top")) +
  coord_sf(expand = FALSE) +
  theme_minimal(base_size = 12) +
    labs(fill = "Total Records", tag = "A") +
  theme(legend.position = c(.08, .39), 
        legend.title = element_text(size = 12,face = "bold"), element_blank(), #element_blank()
        legend.text = element_text(size = 11),
        legend.background = element_rect(fill = "white", color = NA, size = 0.5, linetype = "solid"),
        legend.margin = margin(t = 2, r = 1, b = 2, l = 4),
        plot.title = element_text(hjust = 0.5, face = "bold", size = 14),
        panel.grid.major = element_line(color = "transparent"),
        panel.background = element_rect(fill = "white", color= "NA"),
        plot.margin = margin(10,10,10, 10)) +
  labs(title = "ZOVER Record Locations")
ggsave("figures/fig_2a.jpeg", fig_2a, width = 8, height = 5, dpi = 600)
```

###part b
```{r}
## fix virus name for visualization
top_clean <- top %>% 
    mutate(fix_viruses = case_when(
      fix_viruses == "Lumpy skin disease" ~ "LSD",
      fix_viruses == "African swine fever virus" ~ "ASF",
      fix_viruses == "Colorado tick fever" ~ "CTF",
      fix_viruses == "Crimean-Congo Hemorrhagic fever" ~ "CCHF",
      fix_viruses == "Severe fever with thrombocytopenia syndrome" ~ "SFTS",
      fix_viruses == "Nairobi sheep disease virus" ~ "NSD",
      fix_viruses == "Sawgrass virus" ~ "SGV",
      fix_viruses == "Heartland virus" ~ "HLV",
      fix_viruses == "Bhanja virus" ~ "BHV",
      fix_viruses == "Bourbon virus" ~ "BOV",
      fix_viruses == "Jingmen tick virus" ~ "JTV",
      fix_viruses == "Powassan virus" ~ "PWV",
      fix_viruses == "Tick-borne encephalitis virus" ~ "TBE",
      fix_viruses == "Alkhumra hemorrhagic fever virus" ~ "AHF",
      fix_viruses == "Kyasanur Forest disease virus" ~ "KFD",
      fix_viruses == "Deer tick virus" ~ "DTF",
      fix_viruses == "Omsk hemorrhagic fever virus" ~ "OHF",
      fix_viruses == "Louping ill virus" ~ "LIV",
      TRUE ~ fix_viruses))

## plot it
fig_2b <- top_clean %>% 
  group_by(fix_viruses) %>% 
  tally() %>% 
  mutate(log_n = log(n),
         fix_viruses = reorder(fix_viruses,log_n)) %>% 
  ggplot(aes(y = fix_viruses, x = log_n)) +
  geom_col(col = "#7B50C8", fill = "#CABFF4" , width = .8) + #fill =  "#77278B", color = "#5E1F6A"
  theme_minimal(base_size = 12) +
  theme(axis.text.y = element_text(face = "bold", size = 14),
        axis.title.x = element_text(face = "bold", size = 12),
        plot.title = element_text(face = "bold", size = 14), #, hjust = 0.5
        panel.grid.major.y = element_blank(),
        panel.grid.major.x = element_line(color = "gray90"),
        panel.background = element_rect(fill = "white", color = NA),
        plot.margin = margin(10, 10, 10, 10))  +
  scale_y_discrete(expand = c(0.01, 0.01)) + 
  scale_x_continuous(limits = c(0, 7.6), expand = c(0.01, 0.0)) +
  labs(x = "log(Records in ZOVER)",y = NULL,title = "ZOVER Record Collection",tag = "B")

#ggsave("figures/fig_2b.jpeg", fig_2b, width = 4, height = 5, dpi = 600)

```

###part c
```{r}
## make color pallette
continent_colors <- c("#CB5358", "#587D8C", "#6D9F8E", "#F3B06D")

## wrangle data
virus_order <- top_clean %>% 
  group_by(fix_viruses) %>% 
  tally() %>% 
  arrange(desc(n)) %>% 
  pull(fix_viruses)

# filter
top_filtered <- top_clean %>%
  group_by(admin, fix_viruses, year_agg, continent) %>%
  tally() %>%
  filter(!is.na(continent)) %>%
  mutate(fix_viruses = factor(fix_viruses, levels =  virus_order),
         fix_viruses = fct_rev(fix_viruses),
         year_agg = factor(year_agg, levels = seq(1990, 2020, by = 5)))

## plot base map
fig_2c_base <- ggplot(top_filtered, aes(x = year_agg, y = fix_viruses)) +
  theme_minimal(base_size = 12) +
  theme(axis.text.y = element_text(face = "bold", size = 14),
        axis.text.x = element_text(size = 11.5),
        axis.title = element_text(face = "bold", size = 12),
        plot.title = element_text(face = "bold", size = 14),
        legend.position = "bottom",
        legend.title = element_text(face = "bold", size = 12),
        legend.text = element_text(size = 11),
        legend.key.width = unit(0.3, "cm"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.y = element_line(size = 0.3, color = "grey80"),
        panel.background = element_rect(fill = "white", color = NA),
        plot.margin = margin(10, 10, 10, 10)) +
  labs(x = NULL, y = NULL,color = "Continent",title = "ZOVER Record Timeline",tag = "C") +
  scale_color_manual(values = c("#587D8C", "#6D9F8E", "#F3B06D", "#CB5358")) +
  guides(color = guide_legend(title.position = "top", title.hjust = 0.5))

# make concentric circles
# larger points are plotted first
continents_order <- c("Africa", "Asia", "Europe", "Americas")
sizes <- c(6, 5, 4, 3) 

# write function to plot in order
for (i in seq_along(continents_order)) {
  cont <- continents_order[i]
  fig_2c_base <- fig_2c_base + 
    geom_point(data = filter(top_filtered, continent == cont), 
               aes(color = continent), size = sizes[i],
               show.legend = i == length(continents_order)) # show legend only once
}

## combine base map with concentric circles
fig_2c <- fig_2c_base +
  scale_color_manual(values = continent_colors) +
  scale_y_discrete(limits = levels(top_filtered$fix_viruses))

#ggsave("figures/fig_2c.jpeg", fig_2c, width = 4, height = 5, dpi = 600)

```

Combine b & c
```{r}
fig_2bc <- (fig_2b | (fig_2c + theme(legend.position = "right"))) +
  plot_layout(heights = c(1,1), widths = c(0.5, 0.7),
              guides = "collect")

ggsave("figures/fig_2bc.jpeg", fig_2bc, width = 9, height = 5, dpi = 600)
```

##Figure 3
Variable influence from presence BRT model
```{r}
## color pal
category_colors <- c("Environment" = "#06D6a0",
                     "Wealth" = "#ef476f",
                     "Reporting" = "#ffd166",
                     "Knowledge" = "#118ab2",
                     "Health Care" = "#073b4c",
                     "Exposure" = "#588157")

## assign category to variables
vinf_data_clean <- vinf_data %>% 
  mutate(category = case_when(var == "kopp_climatezone_name" ~ "Environment", 
                              var == "wb_gini_index" ~ "Wealth", 
                              var == "log_sedac_popcounts_sum" ~ "Reporting", 
                              var == "log_fao_pestusecapita_kmp" ~ "Knowledge", 
                              var == "log_wb_landarea_sqkm" ~ "Reporting", 
                              var == "log_wb_livestockprod_index" ~ "Exposure", 
                              var == "ivsa_chapter" ~ "Knowledge", 
                              var == "log_acled_conflictevents_total" ~ "Reporting", 
                              var == "wb_povertyline_pct" ~ "Wealth", 
                              var == "acled_conflictpopexposure_pct" ~ "Reporting",
                              var == "log_glw_livestockdensity_mean" ~ "Exposure",
                              var == "wb_gradvetsforestag_pct" ~ "Knowledge",
                              var == "wb_urbanpop_pct" ~ "Reporting",
                              var == "wb_adultlit_rate" ~ "Knowledge",
                              var == "log_who_reportntd_mean" ~ "Health Care",
                              var == "wb_employedag" ~ "Exposure",
                              var == "wb_gradhealthwelf_pct" ~ "Knowledge",
                              var == "wb_expendhealth_pct" ~ "Health Care",
                              var == "grdi_shdi_mean" ~ "Wealth",
                              var == "wwf_ecoreg_total" ~ "Environment",
                              var == "wb_forested_pct" ~ "Environment",
                              var == "wb_expendedu_pct" ~ "Knowledge",
                              var == "ee_vegdisim_mean" ~ "Environment",
                              var == "wb_accesselect_pct" ~ "Wealth",
                              TRUE ~ NA)) %>%
  mutate(category = factor(category, levels = names(category_colors)))

## make labels nicer
pretty_labels <- rev(c(
  "kopp_climatezone_name" = "Climate Zone",
   "wb_gini_index" = "Gini Index",
   "log_sedac_popcounts_sum" = "Pop. Counts",
   "log_fao_pestusecapita_kmp" = "Pesticide Use",
   "log_wb_landarea_sqkm" = "Land Area",
   "log_wb_livestockprod_index" = "Livestock Production",
   "ivsa_chapter" = "IVSA Chapter Present",
   "log_acled_conflictevents_total" = "Conflict (total events)",
   "wb_povertyline_pct" = "Below Poverty Line",
   "acled_conflictpopexposure_pct" = "Conflict (% of pop.)",
   "log_glw_livestockdensity_mean" = "Livestock Density",
   "wb_gradvetsforestag_pct" = "Vet/Ag/Forest Grads",
   "wb_urbanpop_pct" =  "Urban Pop.",
  "wb_adultlit_rate"  = "Adult Literacy",
   "log_who_reportntd_mean" = "NTD Treatment",
   "wb_employedag" = "Employed in Ag.",
   "wb_gradhealthwelf_pct"  = "Health & Welfare Grads",
   "wb_expendhealth_pct" = "Health Spending",
   "grdi_shdi_mean" = "Social Vulnerability",
   "wwf_ecoreg_total" = "Total Ecoregions",
   "wb_forested_pct"= "Forested",
   "wb_expendedu_pct" = "Education Spending",
   "ee_vegdisim_mean" = "Habitat Fragmentation",
   "wb_accesselect_pct"= "Electricity Access"))

## plot it!
fig_3 <- 
vinf_data_clean %>% 
  arrange(rel.inf) %>% 
  mutate(var = factor(var, levels = unique(var))) %>% 
  ggplot(aes(x = rel.inf, y = var, fill = category)) +
  geom_col(color = "black", width = 0.8) +
  scale_fill_manual(values = category_colors, na.value = "white") +
  scale_x_continuous(expand = c(0,0), limits = c(0, max(vinf_data_clean$rel.inf))) +
  scale_y_discrete(labels = pretty_labels) +
  labs(x = "Relative Influence (%)", y = NULL, fill = "Variable Category") + #, tag = "A"
  theme_minimal(base_size = 12) +
  theme(plot.title = element_text(face = "bold", size = 14),
    axis.title.x = element_text(face = "bold", size = 12),
    axis.text.y = element_text(size = 11, margin = margin(t = 5, b = 5)),
    axis.text.x = element_text(size = 11),
    panel.grid.major.y = element_blank(),
    panel.grid.major.x = element_line(color = "gray90"),
    panel.background = element_rect(fill = "white", color = NA),
    legend.position = c(0.73, 0.23),
    legend.title = element_text(face = "bold", size = 12),
    legend.text = element_text(size = 11, face = "bold"),
    legend.background = element_rect(fill = "white", color = NA),
    plot.margin = margin(10, 10, 10, 10))

ggsave("figures/fig_3.jpeg", fig_3, width = 5, height = 4.5, dpi = 600)
```

##Figure 4
Partial dependence plots from presence BRT model

###part a - write function
```{r}
# step 1. compute PDP values for single model + predictor
pdp_agg <- function(mod, feature) {
  pdep <- plot(mod$mod, feature,
               return.grid = TRUE,
               n.trees = mod$best,
               plot = FALSE,
               continuous.resolution = 200,
               type = "response")
  
  pdep$seed <- mod$seed #unique(mod$roc$seed)
  pdep$predictor <- pdep[[feature]]
  pdep <- pdep[order(pdep$predictor),]
  pdep$rank <- seq_len(nrow(pdep))
  pdep$yhat <- pdep$y
  
  return(pdep)
}

## step 2. make function to plot PDP
pdp_plot <- function(bmods, feature, data, title = NULL, tag = NULL, target = "target_variable") {
  
  # aggregate across BRT models
  agg <- do.call(rbind,lapply(bmods, function(x) pdp_agg(x, feature)))
  
  # determine feature type
  feature_class <- class(data[[feature]])
  
  # set theme style
  base_theme <- theme_minimal(base_size = 12) +
    theme(axis.text.y = element_text(face = "bold", size = 11),
      axis.text.x = element_text(size = 11),
      axis.title.x = element_text(face = "bold", size = 12, margin = margin(t = 5)),
      axis.title.y = element_text(face = "bold", size = 12, margin = margin(r = 5)),
      plot.title = element_text(face = "bold", size = 15),
      panel.grid.major.y = element_blank(),
      panel.grid.major.x = element_blank(), #element_line(color = "gray99"),
      panel.background = element_rect(fill = "white", color = NA),
      plot.margin = margin(10, 10, 10, 10))
  
  if(feature_class %in% c("numeric", "integer")) {
    # continuous predictor
    x_mean <- with(agg, tapply(predictor, rank, mean))
    y_mean <- with(agg, tapply(yhat, rank, mean))
    pmean <- data.frame(predictor = x_mean, yhat = y_mean)
    yrange <- range(c(agg$yhat, pmean$yhat), na.rm = TRUE)
    
    
    # histogram for marginal distributions
    hist_data <- hist(data[[feature]], breaks = 30, plot = FALSE)
    hist_df <- data.frame(mids = hist_data$mids, counts = hist_data$counts)
    hist_df$scaled <- plotrix::rescale(hist_df$counts, yrange)
    p <- ggplot(agg, aes(x = predictor, y = yhat, group = seed)) +
      geom_col(data = hist_df, inherit.aes = FALSE, aes(x = mids, y = scaled), 
               fill = "grey80", width = diff(hist_df$mids)[1], alpha = 0.2) +
     # geom_segment(data = hist_df, inherit.aes = FALSE,
      #             aes(x = mids, xend = mids, y = yrange[1], yend = plotrix::rescale(counts, yrange)),
       #            size = 0.8, color = "grey80", alpha = 0.3) +
      geom_line(size = 1, alpha = 0.25, color = "#CABFF4") + #"grey70") +
      geom_line(data = pmean, aes(x = predictor, y = yhat), inherit.aes = FALSE, size = 1.2, color = "#7B50C8") + #"#77278B"
      labs(x = NULL, y = "Marginal Effect", title = title, tag = tag) + #x = feature
      scale_y_continuous(labels = number_format(accuracy = 0.01)) +
      scale_x_continuous(expand = expansion(mult = c(0.01, 0.01))) +
      base_theme
  } else{
    
    # if predictor is a factor
    y_mean <- with(agg, tapply(yhat, predictor, mean)) 
    pmean <- data.frame(predictor = names(y_mean), yhat = y_mean)
    
    temp <- data
    temp$predictor <- temp[[feature]]
    yrange <- range(c(agg$yhat, pmean$yhat), na.rm = TRUE)
    temp$yhat <- ifelse(temp[[target]] == 1, max(yrange), min(yrange))
    
    p <- ggplot(agg, aes(x = predictor, y = yhat, group = seed)) +
      geom_jitter(size = 1, alpha = 0.25, color ="#CABFF4", width = 0.1) +
      geom_point(data = pmean, aes(x = predictor, y = yhat),
                 inherit.aes = FALSE, size = 2, shape = 15, color = "#7B50C8") + #"#77278b") +
      geom_rug(data = temp, aes(x = predictor, y = yhat),
               inherit.aes = FALSE, sides = "b", position = "jitter",
               color = "grey", alpha = 0.25, na.rm = TRUE) +
      labs(x = NULL, y = "Marginal Effect", title = title, tag = tag) + #x = feature
      scale_y_continuous(limits = c(yrange[1]-0.01, yrange[2] + 0.01),
                         labels = number_format(accuracy = 0.01)) +
      base_theme
  }
  
  return(p)
}

 
```

###part b - apply function
```{r}
## make training dataset
train_data <- brt_presence_results[[1]]$traindata


# make sure things are factors
train_data$kopp_climatezone_name <- as.factor(train_data$kopp_climatezone_name)
train_data$ivsa_chapter <- as.factor(train_data$ivsa_chapter)

## make individual plots
pdp_climate <- pdp_plot(brt_presence_results, "kopp_climatezone_name", data = train_data[!is.na(train_data$kopp_climatezone_name),], title = "Climate Zone", target = "kopp_climatezone_name" )
pdp_ivsa <- pdp_plot(brt_presence_results, "ivsa_chapter", data = train_data[!is.na(train_data$ivsa_chapter),], title = "IVSA Chapter", target = "ivsa_chapter")

pdp_vetagfor <- pdp_plot(brt_presence_results, "wb_gradvetsforestag_pct", data = train_data, title = "Vet/Ag/Forestry Grads")
pdp_gini <- pdp_plot(brt_presence_results, "wb_gini_index", data = train_data, title = "Gini Index")
pdp_pest <- pdp_plot(brt_presence_results, "log_fao_pestusecapita_kmp", data = train_data, title = "Pesticide Use")
pdp_livestock <- pdp_plot(brt_presence_results, "log_wb_livestockprod_index", data = train_data, title = "Livestock Index")
pdp_popcounts <- pdp_plot(brt_presence_results, "log_sedac_popcounts_sum", data = train_data, title = "Population Count")
pdp_conflict <- pdp_plot(brt_presence_results, "acled_conflictpopexposure_pct", data = train_data, title = "% Conflict Exposure")
pdp_poverty <- pdp_plot(brt_presence_results, "wb_povertyline_pct", data = train_data, title = "% Poverty Line")
pdp_land <- pdp_plot(brt_presence_results, "log_wb_landarea_sqkm", data = train_data, title = "Land Area")
pdp_conflicttotal <- pdp_plot(brt_presence_results, "log_acled_conflictevents_total", data = train_data, title = "Tot. Conflict Exposure")
pdp_agemploy <- pdp_plot(brt_presence_results, "wb_employedag", data = train_data, title = "% Ag Employment")
pdp_popurban <- pdp_plot(brt_presence_results, "wb_urbanpop_pct", data = train_data, title = "% Urban Pop.")
pdp_forestarea <- pdp_plot(brt_presence_results, "wb_forested_pct", data = train_data, title = "% Forested Area")
pdp_shdi <- pdp_plot(brt_presence_results, "grdi_shdi_mean", data = train_data, title = "Development Index")
pdp_educationexpend <- pdp_plot(brt_presence_results, "wb_expendedu_pct", data = train_data, title = "Education Expenditures")
pdp_veghetero <- pdp_plot(brt_presence_results, "ee_vegdisim_mean", data = train_data, title = "Habitat Fragmentation")
pdp_healthcarespend <- pdp_plot(brt_presence_results, "wb_expendhealth_pct", data = train_data, title = "Health Expenditures")
pdp_ecortotal <- pdp_plot(brt_presence_results, "wwf_ecoreg_total", data = train_data, title = "Ecoregions Total")
pdp_adultlit <- pdp_plot(brt_presence_results, "wb_adultlit_rate", data = train_data, title = "Adult Literacy Rate Mean")
pdp_electric <- pdp_plot(brt_presence_results, "wb_accesselect_pct", data = train_data, title = "% Electricity Access")
pdp_gradhealth <- pdp_plot(brt_presence_results, "wb_gradhealthwelf_pct", data = train_data, title = "Health & Welfare Grads")
```

###part c - pull out main figures for figure 4
```{r}
pdp_ggarrange <- ggarrange( pdp_gini + rremove("ylab"), pdp_popcounts + rremove("ylab"),
          pdp_livestock + rremove("ylab"), pdp_pest + rremove("ylab") + theme(axis.text.x = element_text(size = 7)), 
          pdp_conflict+ rremove("ylab"),  pdp_poverty + rremove("ylab"), #pdp_vetagfor pdp_poverty
           pdp_ivsa + rremove("ylab"), pdp_climate + rremove("ylab") + theme(axis.text.x = element_text(angle = 90, hjust = 0, size = 7)), 
          #ncol = 4, nrow = 2 
          ncol = 4, nrow = 2)

pdp_ggarrange_an <- annotate_figure(pdp_ggarrange, 
                left = text_grob("Marginal Effect", rot = 90, size = 15, face = "bold"))

ggsave("figures/fig_4_v2.jpeg", pdp_ggarrange_an, width = 10.5, height = 5, dpi = 600) #width = 10


```


#Supplementary Figures

##Figure S1

###part a - citation county by country
```{r}
## set up colors
mypal <- colorRampPalette(rev(brewer.pal(11, "Spectral")))
sc <- scale_fill_gradientn(colors = mypal(8), limits = c(0, 8))

## plot it
fig_s_1a <- top %>% 
  filter(!is.na(iso_a3)) %>% 
  group_by(iso_a3) %>% 
  summarise(mean_citations = mean(pubmedcitations_percounty_total, na.rm = TRUE)) %>% 
  mutate(iso_a3 = case_when(iso_a3 == "France" ~ "FRA",
                            iso_a3 == "Norway" ~ "NOR",
                            TRUE ~ iso_a3)) %>% 
  ggplot(aes(x = log(mean_citations), y = fct_reorder(iso_a3, mean_citations))) +
  geom_col(col = "#7B50C8", fill = "#CABFF4", width = .8) + 
  theme_minimal(base_size = 12) +
  theme(axis.text.y = element_text(face = "bold", size = 11),
        axis.title.x = element_text(face = "bold", size = 12),
        plot.title = element_text(face = "bold", size = 14), #, hjust = 0.5
        panel.grid.major.y = element_blank(),
        panel.grid.major.x = element_line(color = "gray90"),
        panel.background = element_rect(fill = "white", color = NA),
        plot.margin = margin(10, 10, 10, 10))  +
  scale_y_discrete(expand = c(0, 0), guide = guide_axis(check.overlap = TRUE, n.dodge = 2)) + 
  #scale_x_continuous(limits = c(0, 7.6), expand = c(0.01, 0.0)) +
  labs(x = "log(Total Citations)", y = NULL, tags = "A") 

ggsave("figures/fig_s_1a.jpeg",fig_s_1a, width = 4, height = 5, dpi = 600)
```

###part b - citation count by virus
```{r}
top_virus <- top %>% 
group_by(iso_a3,fix_viruses) %>%
  tally() %>% 
  filter(!is.na(iso_a3)) %>% 
  mutate(fix_viruses = case_when(fix_viruses == "Lumpy skin disease" ~ "LSD",
                                        fix_viruses == "African swine fever virus" ~ "ASF",
                                        fix_viruses == "Colorado tick fever" ~ "CTF",
                                        fix_viruses == "Crimean-Congo Hemorrhagic fever" ~ "CCHF",
                                        fix_viruses == "Severe fever with thrombocytopenia syndrome" ~ "SFTS",
                                        fix_viruses == "Nairobi sheep disease virus" ~ "NSD",
                                        fix_viruses == "Sawgrass virus" ~ "SGV",
                                        fix_viruses == "Heartland virus" ~ "HLV",
                                        fix_viruses == "Bhanja virus" ~ "BHV",
                                        fix_viruses == "Bourbon virus" ~ "BOV",
                                        fix_viruses == "Jingmen tick virus" ~ "JTV",
                                        fix_viruses == "Powassan virus" ~ "PWV",
                                        fix_viruses == "Tick-borne encephalitis virus" ~ "TBE",
                                        fix_viruses == "Alkhumra hemorrhagic fever virus" ~ "AHF",
                                        fix_viruses == "Kyasanur Forest disease virus" ~ "KFD",
                                        fix_viruses == "Deer tick virus" ~ "DTF",
                                        fix_viruses == "Omsk hemorrhagic fever virus" ~ "OHF",
                                        fix_viruses == "Louping ill virus" ~ "LIV",
                                        TRUE ~ fix_viruses),
         iso_a3 = case_when(iso_a3 == "France" ~ "FRA",
                            iso_a3 == "Norway" ~ "NOR",
                            TRUE ~ iso_a3))  


fig_s_1b <- top_virus  %>% 
  ggplot(aes(x = fix_viruses, y = fct_rev(factor(iso_a3)), fill = log(n+1))) +
  geom_tile() +
  labs(x = NULL, y = NULL, fill = "log(Total Citations)", tags = "B") +
  scale_y_discrete(expand = c(0, 0), guide = guide_axis(check.overlap = TRUE, n.dodge = 2)) +
  sc +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, face = "bold", size = 12),
        axis.text.y = element_text(face = "bold", size = 11),
        panel.background = element_rect(fill = "#f9f9f9", color = NA),
        #plot.background = element_rect(fill = "#f9f9f9", color = NA),
    legend.position = "bottom", #"right",
    legend.margin = margin(t = -5),
    legend.box.margin = margin(t = -5),
    legend.title = element_text(face = "bold", vjust = .9)) 

ggsave("figures/fig_s_1b.jpeg",fig_s_1b, width = 4, height = 5, dpi = 600)

```

combine them
```{r}
fig_s_1ab <- ggarrange(fig_s_1a, fig_s_1b)

ggsave("figures/fig_s_1ab.jpeg",fig_s_1ab, width = 8, height = 10, dpi = 600)

```

##Figure S2
```{r}

## write function to plot PDP
pdp_plot_ylimslim <- function(bmods, feature, data, title = NULL, tag = NULL, target = "target_variable") {
  
  # aggregate across BRT models
  agg <- do.call(rbind,lapply(bmods, function(x) pdp_agg(x, feature)))
  
  # determine feature type
  feature_class <- class(data[[feature]])
  
  # set theme style
  base_theme <- theme_minimal(base_size = 12) +
    theme(axis.text.y = element_text(face = "bold", size = 11),
      axis.text.x = element_text(size = 11),
      axis.title.x = element_text(face = "bold", size = 12, margin = margin(t = 5)),
      axis.title.y = element_text(face = "bold", size = 12, margin = margin(r = 5)),
      plot.title = element_text(face = "bold", size = 15),
      panel.grid.major.y = element_blank(),
      panel.grid.major.x = element_blank(), #element_line(color = "gray99"),
      panel.background = element_rect(fill = "white", color = NA),
      plot.margin = margin(10, 10, 10, 10))
  
  if(feature_class %in% c("numeric", "integer")) {
    # continuous predictor
    x_mean <- with(agg, tapply(predictor, rank, mean))
    y_mean <- with(agg, tapply(yhat, rank, mean))
    pmean <- data.frame(predictor = x_mean, yhat = y_mean)
    #yrange <- range(c(agg$yhat, pmean$yhat), na.rm = TRUE)
    
      # automatically calculate yrange with buffer
  y_min <- min(c(agg$yhat, pmean$yhat), na.rm = TRUE)
  y_max <- max(c(agg$yhat, pmean$yhat), na.rm = TRUE)
  buffer <- (y_max - y_min) * 0.05  # 5% padding
  yrange <- c(y_min - buffer, y_max + buffer)
  
    
    # histogram for marginal distributions
    hist_data <- hist(data[[feature]], breaks = 30, plot = FALSE)
    hist_df <- data.frame(mids = hist_data$mids, counts = hist_data$counts)
    hist_df$scaled <- plotrix::rescale(hist_df$counts, yrange)
    p <- ggplot(agg, aes(x = predictor, y = yhat, group = seed)) +
      geom_col(data = hist_df, inherit.aes = FALSE, aes(x = mids, y = scaled), 
               fill = "grey80", width = diff(hist_df$mids)[1], alpha = 0.2) +
     # geom_segment(data = hist_df, inherit.aes = FALSE,
      #             aes(x = mids, xend = mids, y = yrange[1], yend = plotrix::rescale(counts, yrange)),
       #            size = 0.8, color = "grey80", alpha = 0.3) +
      geom_line(size = 1, alpha = 0.25, color = "#CABFF4") + #"grey70") +
      geom_line(data = pmean, aes(x = predictor, y = yhat), inherit.aes = FALSE, size = 1.2, color = "#7B50C8") + #"#77278B"
      labs(x = NULL, y = "Marginal Effect", title = title, tag = tag) + #x = feature
      scale_y_continuous(labels = number_format(accuracy = 0.01),limits = yrange) +
      scale_x_continuous(expand = expansion(mult = c(0.01, 0.01))) +
      base_theme
  } else{
    
    # if predictor is a factor
    y_mean <- with(agg, tapply(yhat, predictor, mean)) 
    pmean <- data.frame(predictor = names(y_mean), yhat = y_mean)
    
    temp <- data
    temp$predictor <- temp[[feature]]
    yrange <- range(c(agg$yhat, pmean$yhat), na.rm = TRUE)
    temp$yhat <- ifelse(temp[[target]] == 1, max(yrange), min(yrange))
    
    p <- ggplot(agg, aes(x = predictor, y = yhat, group = seed)) +
      geom_jitter(size = 1, alpha = 0.25, color ="#CABFF4", width = 0.1) +
      geom_point(data = pmean, aes(x = predictor, y = yhat),
                 inherit.aes = FALSE, size = 2, shape = 15, color = "#7B50C8") + #"#77278b") +
      geom_rug(data = temp, aes(x = predictor, y = yhat),
               inherit.aes = FALSE, sides = "b", position = "jitter",
               color = "grey", alpha = 0.25, na.rm = TRUE) +
      labs(x = NULL, y = "Marginal Effect", title = title, tag = tag) + #x = feature
      scale_y_continuous(limits = c(yrange[1]-0.01, yrange[2] + 0.01),
                         labels = number_format(accuracy = 0.01)) +
      base_theme
  }
  
  return(p)
}




## make individual plots
slim_pdp_climate <- pdp_plot_ylimslim(brt_presence_results, "kopp_climatezone_name", data = train_data[!is.na(train_data$kopp_climatezone_name),], title = "Climate Zone", target = "kopp_climatezone_name" )
slim_pdp_ivsa <- pdp_plot_ylimslim(brt_presence_results, "ivsa_chapter", data = train_data[!is.na(train_data$ivsa_chapter),], title = "IVSA Chapter", target = "ivsa_chapter")

slim_pdp_vetagfor <- pdp_plot_ylimslim(brt_presence_results, "wb_gradvetsforestag_pct", data = train_data, title = "Vet/Ag/Forestry Grads")
slim_pdp_gini <- pdp_plot_ylimslim(brt_presence_results, "wb_gini_index", data = train_data, title = "Gini Index")
slim_pdp_pest <- pdp_plot_ylimslim(brt_presence_results, "log_fao_pestusecapita_kmp", data = train_data, title = "Pesticide Use")
slim_pdp_livestock <- pdp_plot_ylimslim(brt_presence_results, "log_wb_livestockprod_index", data = train_data, title = "Livestock Index")
slim_pdp_popcounts <- pdp_plot_ylimslim(brt_presence_results, "log_sedac_popcounts_sum", data = train_data, title = "Population Count")
slim_pdp_conflict <- pdp_plot_ylimslim(brt_presence_results, "acled_conflictpopexposure_pct", data = train_data, title = "% Conflict Exposure")
slim_pdp_poverty <- pdp_plot_ylimslim(brt_presence_results, "wb_povertyline_pct", data = train_data, title = "% Poverty Line")
slim_pdp_land <- pdp_plot_ylimslim(brt_presence_results, "log_wb_landarea_sqkm", data = train_data, title = "Land Area")
slim_pdp_conflicttotal <- pdp_plot_ylimslim(brt_presence_results, "log_acled_conflictevents_total", data = train_data, title = "Tot. Conflict Exposure")
slim_pdp_agemploy <- pdp_plot_ylimslim(brt_presence_results, "wb_employedag", data = train_data, title = "% Ag Employment")
slim_pdp_popurban <- pdp_plot_ylimslim(brt_presence_results, "wb_urbanpop_pct", data = train_data, title = "% Urban Pop.")
slim_pdp_forestarea <- pdp_plot_ylimslim(brt_presence_results, "wb_forested_pct", data = train_data, title = "% Forested Area")
slim_pdp_shdi <- pdp_plot_ylimslim(brt_presence_results, "grdi_shdi_mean", data = train_data, title = "Development Index")
slim_pdp_educationexpend <- pdp_plot_ylimslim(brt_presence_results, "wb_expendedu_pct", data = train_data, title = "Education Expenditures")
slim_pdp_veghetero <- pdp_plot_ylimslim(brt_presence_results, "ee_vegdisim_mean", data = train_data, title = "Habitat Fragmentation")
slim_pdp_healthcarespend <- pdp_plot_ylimslim(brt_presence_results, "wb_expendhealth_pct", data = train_data, title = "Health Expenditures")
slim_pdp_ecortotal <- pdp_plot_ylimslim(brt_presence_results, "wwf_ecoreg_total", data = train_data, title = "Ecoregions Total")
slim_pdp_adultlit <- pdp_plot_ylimslim(brt_presence_results, "wb_adultlit_rate", data = train_data, title = "Adult Literacy Rate Mean")
slim_pdp_electric <- pdp_plot_ylimslim(brt_presence_results, "wb_accesselect_pct", data = train_data, title = "% Electricity Access")
slim_pdp_gradhealth <- pdp_plot_ylimslim(brt_presence_results, "wb_gradhealthwelf_pct", data = train_data, title = "Health & Welfare Grads")


## arrange plots
slim_pdp_ggarrange <- ggarrange(slim_pdp_climate + rremove("ylab") + theme(axis.text.x = element_text(angle = 90, hjust = 0, size = 7)),
          slim_pdp_ivsa + rremove("ylab"), 
          slim_pdp_vetagfor + rremove("ylab"),
          slim_pdp_gini + rremove("ylab"),
          slim_pdp_pest + rremove("ylab"),
          slim_pdp_livestock + rremove("ylab"),
          slim_pdp_popcounts + rremove("ylab"),
          slim_pdp_conflict + rremove("ylab"),
          slim_pdp_poverty + rremove("ylab"),
          slim_pdp_land + rremove("ylab"),
          slim_pdp_conflicttotal + rremove("ylab"),
          slim_pdp_agemploy + rremove("ylab"),
          slim_pdp_popurban + rremove("ylab"),
          slim_pdp_forestarea + rremove("ylab"), 
          slim_pdp_shdi + rremove("ylab"),
          slim_pdp_educationexpend + rremove("ylab"),
          slim_pdp_veghetero + rremove("ylab"),
          slim_pdp_ecortotal + rremove("ylab"),
          slim_pdp_adultlit + rremove("ylab"),
          slim_pdp_electric + rremove("ylab"),
          slim_pdp_healthcarespend + rremove("ylab"),
          slim_pdp_gradhealth + rremove("ylab"),
          ncol = 6, nrow = 4)

## annotate figure
slim_pdp_ggarrange_an <- annotate_figure(slim_pdp_ggarrange, 
                left = text_grob("Marginal Effect", rot = 90, size = 15, face = "bold"))


ggsave("figures/fig_s_2.jpeg", slim_pdp_ggarrange_an, width = 15, height = 10, dpi = 600) 
```

##Figure S3
```{r}
### format data
## factor to make plotting easier
hsearch_results$shrinkage <- factor(hsearch_results$shrinkage)
lvl <-  rev(sort(unique(hsearch_results$shrinkage)))
hsearch_results$shrinkage <- factor(hsearch_results$shrinkage, levels = lvl); rm(lvl)
hsearch_results$interaction.depth <- factor(hsearch_results$interaction.depth)

## convert from wide to long
hsearch_results_long <- gather(hsearch_results, measure, value, testAUC:spec)

# chnage factors
hsearch_results_long$measure <- factor(hsearch_results_long$measure,
                               levels = c("testAUC", "sen", "spec"),
                               labels = c("test AUC", "sensitivity", "specificity"))

### plot data
#### part A
fig_s_3a <- hsearch_results_long %>% 
  ggplot(aes(x = shrinkage, y = value, color = interaction.depth, fill = interaction.depth)) +
  geom_boxplot(alpha = 0.2, position = position_dodge(width = 0.75)) +
  geom_point(alpha = 0.7, size = 2,position = position_jitterdodge(jitter.width = 0.2, dodge.width = 0.75)) +
  theme_minimal(base_size = 12) +
  theme( panel.grid.major.y = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.text = element_text(size = 10, face = "bold"),
    axis.title.x = element_text(size = 12, face = "bold", margin = margin(t = 10)),
    axis.title.y = element_text(size = 12, face = "bold", margin = margin(r = 10)),
    legend.title = element_text(face = "bold"),
    legend.text = element_text(size = 10),
    legend.position = "bottom",
    plot.margin = margin(10, 10, 10, 10)) +
  labs(x = "Learning Rate", y = "Model Performance", color = "Interaction Depth", tag = "A") +
  scale_y_continuous(n.breaks = 4) +
  scale_color_manual(values = c("#587D8C",  "#F3B06D", "#CB5358")) + #"#6D9F8E",
  scale_fill_manual(values = c("#587D8C",  "#F3B06D", "#CB5358")) +
  guides(fill = FALSE)

ggsave("figures/fig_s_3a.jpeg",fig_s_3a, width = 4, height = 4, dpi = 600) 

#### part B
# Summarize mean test AUC for each hyperparameter combination
heat_data <- hsearch_results %>% 
  group_by(shrinkage, interaction.depth) %>%
  summarise(mean_testAUC = mean(testAUC, na.rm = TRUE),
            sd_testAUC = sd(testAUC, na.rm = TRUE),
            .groups = 'drop')

# Plot heatmap
fig_s_3b <- ggplot(heat_data, aes(x = interaction.depth, y = shrinkage, fill = mean_testAUC)) +
  geom_tile(color = "white") +
  geom_text(aes(label = round(mean_testAUC, 4)), color = "black", size = 4) +
  scale_fill_viridis_c(name = "Mean Test AUC", option = "C", direction = -1) +
  labs(tag = "B",
    title = "Mean Test AUC by Hyperparameter Combination",
    x = "Interaction Depth",
    y = "Learning Rate (Shrinkage)"
  ) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "right",
    panel.grid = element_blank(),
    plot.title = element_text(face = "bold", hjust = 0.5),
    axis.text = element_text(size = 10, face = "bold"),
    axis.title.x = element_text(size = 12, face = "bold", margin = margin(t = 10)),
    axis.title.y = element_text(size = 12, face = "bold", margin = margin(r = 10)),
    legend.title = element_text(face = "bold"),
    legend.text = element_text(size = 10))

### combine them
fig_s_3ab <- ggarrange(fig_s_3a, fig_s_3b, ncol = 1, align = "hv")

ggsave(fig_s_3ab, file = "figures/fig_s_3ab.jpeg", width = 8, height = 8, dpi = 600)


```

##Figure S4
```{r}

## set up colors
my_gradient <- colorRampPalette(c("#ECE9FC", "#CABFF4", "#A495EB", "#8169E4", "#7B50C8", "#783CA9", "#77278B"))(100)

## call in spatial data
world_matrix_data <- left_join(worldmap_slim , brt_matrix_data%>% dplyr::select(-admin), by = "iso_a3")

## make function to make consistent maps
covariate_map <- function(data, var, var_label = "", my_gradient = NULL, limits = NULL, 
                          legend_position = c(0.08, 0.39),log_transform = FALSE, log_base = 10) {
  
  var_sym <- rlang::ensym(var)
  var_name <- as.character(var_sym)
  
  # log transform
  if (log_transform) {
    data[[var_name]] <- log(data[[var_name]] + 1, base = log_base)
  }
  
  if(is.null(my_gradient)) {
    my_gradient = c("#ECE9FC", "#CABFF4", "#A495EB", "#8169E4", "#7B50C8", "#783CA9", "#77278B")
  }
  
  if(is.null(limits)) {
    limits <- range(data[[as.character(var_sym)]], na.rm = TRUE)
  }
  
  ggplot() +
    geom_sf(data = data, aes(fill = !!var_sym), color = NA) +
    geom_sf(data = world, fill = NA, color = "black", size = 0.2) +
    scale_fill_gradientn(colors = my_gradient, limits = limits, 
                         labels = scales::label_number(accuracy = 0.01), na.value = "#FCFBFF",
                         guide = guide_colorbar(direction = "vertical", barwidth = unit(0.4, "cm"), barheight = unit(4, "cm"),
                                              label.hjust = 0, label.position = "right", title.position = "top")) +
    coord_sf(expand = FALSE) +
    theme_void(base_size = 12) +
    labs(fill = var_label) +
      theme(legend.position = legend_position, 
        legend.title = element_text(size = 12,face = "bold"), element_blank(), 
        legend.text = element_text(size = 11),
        legend.background = element_rect(fill = "white", color = NA, size = 0.5, linetype = "solid"),
        legend.margin = margin(t = 2, r = 1, b = 2, l = 4),
        plot.title = element_text(face = "bold", size = 14),
        panel.grid.major = element_line(color = "transparent"),
        panel.background = element_rect(fill = "white", color= "NA"),
        plot.margin = margin(10,10,10, 10)) 
}

## apply function to covariates
map_gini <- covariate_map(world_matrix_data, var = "wb_gini_index", var_label = "Gini Index") + ggtitle("Gini Index")
map_popcounts <- covariate_map(world_matrix_data, var = "log_sedac_popcounts_sum", var_label = "Counts") + ggtitle("log(Human Population Counts)")
map_pesti <- covariate_map(world_matrix_data, var = "log_fao_pestusecapita_kmp", var_label = "Per Capita", log_transform = TRUE) + ggtitle("log(Pesticide Use per Capita")
map_livestockprod <- covariate_map(world_matrix_data, var = "log_wb_livestockprod_index", var_label = "Index") + ggtitle("Livestock Production Index")
map_conflicttotal <- covariate_map(world_matrix_data, var = "log_acled_conflictevents_total", var_label = "Total") + ggtitle("log(Conflict Exposure Total Events)")
map_poverty <- covariate_map(world_matrix_data, var = "wb_povertyline_pct", var_label = "%") + ggtitle("% of Population Below National Poverty Line")
# need to do ISVA chapter
map_conflictperc <- covariate_map(world_matrix_data, var = "acled_conflictpopexposure_pct", var_label = "%") + ggtitle("Proportion of Population Exposed to Conflict")
map_agvetfor <- covariate_map(world_matrix_data, var = "wb_gradvetsforestag_pct", var_label = "%") + ggtitle("% of Graduates from Ag, Vet, Forestry Tertiary Schools")
map_agemployed <- covariate_map(world_matrix_data, var = "wb_employedag", var_label = "%") + ggtitle("% of Population Employed in Ag")
map_ivsa <- ggplot() +
    geom_sf(data = world_matrix_data, aes(fill = as.factor(ivsa_chapter)), color = NA) +
    geom_sf(data = world, fill = NA, color = "black", size = 0.2) +
    scale_fill_manual(values = c("0" =  "#77278B", "1" = "#ECE9FC"), na.value = "#FCFBFF",labels = c("Absent", "Present",""),
                      guide = guide_legend(direction = "vertical", keywidth = unit(0.4, "cm"), keyheight = unit(0.4, "cm"),
                                              label.hjust = 0, label.position = "right", title.position = "top")) +
    coord_sf(expand = FALSE) +
    theme_void(base_size = 12) +
    labs(fill = "IVSA") +
      theme(legend.position = c(0.08, 0.39), 
        legend.title = element_text(size = 12,face = "bold"), element_blank(), 
        legend.text = element_text(size = 11),
        legend.background = element_rect(fill = "white", color = NA, size = 0.5, linetype = "solid"),
        legend.margin = margin(t = 2, r = 1, b = 2, l = 4),
        plot.title = element_text(face = "bold", size = 14),
        panel.grid.major = element_line(color = "transparent"),
        panel.background = element_rect(fill = "white", color= "NA"),
        plot.margin = margin(10,10,10, 10))  +
  ggtitle("IVSA Chapter Status")

## combine figures
fig_s_4a <- ggarrange(map_gini, map_poverty, map_popcounts, map_pesti, 
          ncol = 2, nrow = 2, align = "hv")

ggsave(fig_s_4a, file = "figures/fig_s_4a.jpeg", width = 11, height = 7, dpi = 300) 

fig_s_4b <- ggarrange(map_livestockprod, map_ivsa, map_conflictperc,  map_agemployed, 
          ncol = 2, nrow = 2, align = "hv")
ggsave(fig_s_4b, file = "figures/fig_s_4b.jpeg", width = 11, height = 7, dpi = 300) 


```

##Figure S5
```{r}
## part A
fig_s_5a <- ggplot() +
    geom_sf(data = world_matrix_data, aes(fill = tbv_total), color = NA) +
    geom_sf(data = world, fill = NA, color = "black", size = 0.2) +
    scale_fill_gradientn(colors = my_gradient, limits = c(min(world_matrix_data$tbv_total, na.rm = TRUE), max(world_matrix_data$tbv_total, na.rm = TRUE)) ,
                         na.value = "#FCFBFF",
                         guide = guide_colorbar(direction = "vertical", barwidth = unit(0.4, "cm"), barheight = unit(4, "cm"),
                                              label.hjust = 0, label.position = "right", title.position = "top")) +
    coord_sf(expand = FALSE) +
    theme_void(base_size = 12) +
    labs(fill = "Total") +
      theme(legend.position = c(0.08, 0.39), 
        legend.title = element_text(size = 12,face = "bold"), element_blank(), 
        legend.text = element_text(size = 11),
        legend.background = element_rect(fill = "white", color = NA, size = 0.5, linetype = "solid"),
        legend.margin = margin(t = 2, r = 1, b = 2, l = 4),
        plot.title = element_text(face = "bold", size = 14),
        panel.grid.major = element_line(color = "transparent"),
        panel.background = element_rect(fill = "white", color= "NA"),
        plot.margin = margin(10,10,10, 10)) +
  ggtitle("Total Records in ZOVER for select TBVs")

ggsave("figures/fig_s_5a.jpeg",fig_s_5a, width = 4, height = 5, dpi = 600)

## part B
full_tbv_record <- all_tbvs %>% 
  group_by(sampling_country) %>% 
  tally() %>% 
  rename(total_viruses_all = n,
         admin = sampling_country) 


full_tbv_record_sf <- world %>% 
  left_join(full_tbv_record, by = "admin")

fig_s_5b <- ggplot() +
    geom_sf(data = full_tbv_record_sf, aes(fill = total_viruses_all), color = NA) +
    geom_sf(data = world_matrix_data, fill = NA, color = "black", size = 0.2) +
    scale_fill_gradientn(colors = my_gradient, limits = c(min(full_tbv_record_sf$total_viruses_all, na.rm = TRUE), max(full_tbv_record_sf$total_viruses_all, na.rm = TRUE)) ,
                         na.value = "#FCFBFF",
                         guide = guide_colorbar(direction = "vertical", barwidth = unit(0.4, "cm"), barheight = unit(4, "cm"),
                                              label.hjust = 0, label.position = "right", title.position = "top")) +
    coord_sf(expand = FALSE) +
    theme_void(base_size = 12) +
    labs(fill = "Total") +
      theme(legend.position = c(0.08, 0.39), 
        legend.title = element_text(size = 12,face = "bold"), element_blank(), 
        legend.text = element_text(size = 11),
        legend.background = element_rect(fill = "white", color = NA, size = 0.5, linetype = "solid"),
        legend.margin = margin(t = 2, r = 1, b = 2, l = 4),
        plot.title = element_text(face = "bold", size = 14),
        panel.grid.major = element_line(color = "transparent"),
        panel.background = element_rect(fill = "white", color= "NA"),
        plot.margin = margin(10,10,10, 10)) +
  ggtitle("Total Records in ZOVER for all TBVs")

ggsave("figures/fig_s_5b.jpeg",fig_s_5b, width = 4, height = 5, dpi = 600)

### combine figures
fig_s_5ab <- ggarrange(fig_s_5a, fig_s_5b, ncol = 1, nrow = 2)
ggsave("figures/fig_s_5ab.jpeg", fig_s_5ab, width = 8, height = 9, dpi = 600)
```

##Figure S6
```{r}
# pull model info
mod <- brt_presence_results[[1]]$mod      # Access the first model
best.iter <- brt_presence_results[[1]]$best 

# pull result per zone
pdp_data <- plot(mod, i.var = "kopp_climatezone_name", n.trees = best.iter, return.grid = TRUE)

# reorder from most to least important zone
pdp_data$kopp_climatezone_name <- reorder(pdp_data$kopp_climatezone_name, pdp_data$y)
 
# make legend to connect zones to names
koepp_legend <- data.frame(id = seq(1, 32, 1),
                            climate = c('Af', 'Am', 'As', 'Aw', 'BSh', 'BSk', 'BWh', 'BWk', 
                                        'Cfa', 'Cfb', 'Cfc', 'Csa', 'Csb', 'Csc', 'Cwa', 'Cwb', 
                                        'Cwc', 'Dfa', 'Dfb', 'Dfc', 'Dfd', 'Dsa', 'Dsb', 'Dsc', 
                                        'Dsd', 'Dwa', 'Dwb', 'Dwc', 'Dwd', 'EF', 'ET', 'Ocean'),
                            stringsAsFactors = FALSE)
 # spell out the full names like on wikipedia

koepp_legend$climate_full_name <- c(
   "Tropical rainforest climate",                                 # Af
   "Tropical monsoon climate",                                    # Am
   "Tropical savanna climate with dry summer",                    # As
   "Tropical savanna climate with dry winter",                    # Aw
   "Dry Hot semi-arid climate (Steppe, hot)",                         # BSh
   "Dry Cold semi-arid climate (Steppe, cold)",                       # BSk
   "Dry Hot desert climate",                                          # BWh
   "Dry Cold desert climate",                                         # BWk
   "Temperate Humid subtropical climate, no dry season, hot summer",       # Cfa
   "Temperate Oceanic climate, no dry season, warm summer",                 # Cfb
   "Temperate Subpolar oceanic climate, no dry season, cool summer",       # Cfc
   "Temperate Mediterranean climate, dry hot summer",                       # Csa
   "Temperate Mediterranean climate, dry warm summer",                      # Csb
   "Temperate Mediterranean climate, dry cool summer",                      # Csc
   "Temperate Monsoon-influenced humid subtropical climate, dry winter, hot summer", # Cwa
   "Temperate Subtropical highland climate or monsoon-influenced, dry winter, warm summer", # Cwb
   "Temperate Subtropical highland climate or monsoon-influenced, dry winter, cool summer", # Cwc
   "Continental Humid continental climate, no dry season, hot summer",       # Dfa
   "Continental Humid continental climate, no dry season, warm summer",      # Dfb
   "Continental Subarctic climate, no dry season, cool summer",              # Dfc
   "Continental Subarctic climate, no dry season, very cold winter",         # Dfd
   "Continental Mediterranean climate, dry hot summer",          # Dsa
   "Continental Mediterranean climate, dry warm summer",         # Dsb
   "Continental Mediterranean climate, dry cool summer",         # Dsc
   "Continental Mediterranean climate, dry very cold winter",    # Dsd
   "Continental Monsoon-influenced humid continental climate, dry winter, hot summer", # Dwa
   "Continental Monsoon-influenced humid continental climate, dry winter, warm summer", # Dwb
   "Continental Monsoon-influenced subarctic climate, dry winter, cool summer", # Dwc
   "Continental Monsoon-influenced subarctic climate, dry winter, very cold winter", # Dwd
   "Polar Ice cap climate",                                             # EF
   "Polar Tundra climate",                                              # ET
   "Oceanic climate (general label)")                            # Ocean


pdp_zone <- pdp_data %>% 
  left_join(koepp_legend %>% rename(kopp_climatezone_name = climate), by = "kopp_climatezone_name")


pdp_zone$climate_full_name <- reorder(pdp_zone$climate_full_name, pdp_data$y)

# plot it

fig_s_6 <- pdp_zone %>% 
  ggplot(aes(x = climate_full_name, y = y)) +
  geom_point(color =  "#77278B", size = 2) +
  labs(title = "Climate Classification", x = "", y = "Partial dependence") +
  theme_minimal() +
  coord_flip() +
  theme(title = element_text(face = "bold"))

ggsave("figures/fig_s_6.jpeg", fig_s_6, width = 8, height = 8, dpi = 600)
```
